---
title: "Descriptive Information"
format: 
  html:
    embed-resources: true
    fig-width: 8
    fig-height: 8
    code-link: true
    code-fold: true
    code-tools: true
    df-print: paged
    toc: true
    grid:
      body-width: 900px
editor: source
editor_options: 
  chunk_output_type: console
---

# Load initial script

```{r}
#| output: false

source("0_load_data.R")
```

# Table describing each variable 

```{r}

# Function to map variable prefix to study wave
get_study_wave = function(var_name) {
  prefix = substr(var_name, 1, 1)
  wave_map = c(
    "a" = "1st Contact",
    "b" = "2 Year",
    "c" = "3 Year",
    "d" = "4 Year",
    "e" = "In Home",
    "g" = "7 Year",
    "h" = "8 Year",
    "i" = "9 Year",
    "j" = "10 Year",
    "l" = "12 Year",
    "n" = "14 Year",
    "p" = "16 Year",
    "r" = "18 Year",
    "u" = "21 Year",
    "z" = "26 Year"
  )
  return(wave_map[prefix])
}

# Function to clean description labels
clean_description = function(x) {
  x %>%
    # Remove range info at end (e.g., ", 0-22", ", 1-9", ", 4-11")
    str_remove(", \\d+-\\d+$") %>%
    # Remove "see value labels"
    str_remove(", see value labels") %>%
    str_remove("see value labels") %>%
    # Spell out abbreviations inline
    str_replace("SOC employment", "Standard Occupational Classification employment") %>%
    str_replace("MHQ\\)", "Mental Health Questionnaire)") %>%
    str_replace("SLQ\\)", "questionnaire)") %>%
    str_replace("G-game", "G-game (general cognitive ability)") %>%
    str_replace_all(" qnr", " questionnaire") %>%
    # Clean up extra whitespace

    str_squish()
}

v_rq1x = 
  data.frame(
    `Short Label` = rq1x_labels_clean,
    Description = rq1x_labels,
    `Teds Code` = ifelse(rq1x %in% colnames(df0), rq1x,paste0(rq1x,"*")),
    `Range or Level` = sapply(rq1x, function(var) {
      if (class(df[[var]]) == "numeric") {
        paste0(round(min(df[[var]], na.rm = TRUE), 2), " — ", round(max(df[[var]], na.rm = TRUE), 2))
      } else if (is.factor(df[[var]])) {
        factor_levels = levels(df[[var]])
        paste(c(paste0(factor_levels[1],"*"), factor_levels[-1]), collapse = ", ")
      } else {
        paste(unique(df[[var]]), collapse = ", ")
      }
    }),
    N = sapply(rq1x, function(var) {
      sum(!is.na(df[[var]]))
    }),
    `Study Wave` = sapply(rq1x, get_study_wave)
  )

v_rq1x$Description = str_remove(v_rq1x$Description,"\\(1st.*")
v_rq1x$Description = clean_description(v_rq1x$Description)

# Manual edit for cens01pop98density study wave
v_rq1x$`Study.Wave`[v_rq1x$`Teds.Code` == "cens01pop98density"] = "1st Contact"
v_rq1x$`Study.Wave`[v_rq1x$`Teds.Code` == "pollution1998pca"] = "1st Contact"

# Clean uniform descriptions for participation indicator variables
rq1y_twin_descriptions = c(
  "Parent-report twin booklet data present",
  "Web test data present",
  "Self-report questionnaire data present",
  "Self-report behaviour booklet data present",
  "Self-report questionnaire data present",
  "TEDS21 phase 1 self-report questionnaire data present",
  "TEDS26 Mental Health Questionnaire data present",
  "CATSLife web test data present"
)

v_rq1y =
  data.frame(
    `Short Label` = rq1y_twin_labels_clean,
    Description = rq1y_twin_descriptions,
    `Teds Code` = ifelse(rq1y_twin1 %in% colnames(df0), rq1y_twin1, paste0(rq1y_twin1, "*")),
    `Range or Level` = sapply(rq1y_twin1, function(var) {
      if (class(df[[var]]) == "numeric") {
        paste0(round(min(df[[var]], na.rm = TRUE), 2), " — ", round(max(df[[var]], na.rm = TRUE), 2))
      } else if (is.factor(df[[var]])) {
        factor_levels = levels(df[[var]])
        paste(c(paste0(factor_levels[1], "*"), factor_levels[-1]), collapse = ", ")
      } else {
        paste(unique(df[[var]]), collapse = ", ")
      }
    }),
    N = sapply(rq1y_twin1, function(var) {
      sum(!is.na(df[[var]]))
    }),
    `Study Wave` = sapply(rq1y_twin, get_study_wave)
  )

v_rq2 = data.frame(
  `Short Label` = rq2y_labels_short,
  Description   = rq2y_labels,
  `Teds Code` = ifelse(rq2y %in% colnames(df0), rq2y,paste0(rq2y,"*")),
  `Range or Level` = sapply(rq2y, function(var) {
    if (class(df[[var]]) == "numeric") {
      paste0(round(min(df[[var]], na.rm = TRUE), 2), " — ", round(max(df[[var]], na.rm = TRUE), 2))
    } else if (is.factor(df[[var]])) {
      factor_levels = levels(df[[var]])
      paste(c(paste0(factor_levels[1],"*"), factor_levels[-1]), collapse = ", ")
    } else {
      paste(unique(df[[var]]), collapse = ", ")
    }
  }),
  N = sapply(rq2y, function(var) {
    sum(!is.na(df[[var]]))
  }),
  `Study Wave` = sapply(rq2y, get_study_wave)
)

v_rq2$Description = str_remove(v_rq2$Description,"\\(2.*")

v_rq3 = data.frame(
  `Short Label` = rq6y_labels,
  Description   = clean_description(var_to_label(rq6y)),
  `Teds Code` = ifelse(rq6y %in% colnames(df0), rq6y, paste0(rq6y, "*")),
  `Range or Level` = sapply(rq6y, function(var) {
    if (class(df[[var]]) == "numeric") {
      paste0(round(min(df[[var]], na.rm = TRUE), 2), " — ", round(max(df[[var]], na.rm = TRUE), 2))
    } else if (is.factor(df[[var]])) {
      factor_levels = levels(df[[var]])
      paste(c(paste0(factor_levels[1], "*"), factor_levels[-1]), collapse = ", ")
    } else {
      paste(unique(df[[var]]), collapse = ", ")
    }
  }),
  N = sapply(rq6y, function(var) {
    sum(!is.na(df[[var]]))
  }),
  `Study Wave` = sapply(rq6y, get_study_wave)
)

v_rq5 = data.frame(
  `Short Label` = rq5y_labels_short,
  Description   = clean_description(var_to_label(rq5y)),
  `Teds Code` = ifelse(rq5y %in% colnames(df0), rq5y, paste0(rq5y, "*")),
  `Range or Level` = sapply(rq5y, function(var) {
    if (class(df[[var]]) == "numeric") {
      paste0(round(min(df[[var]], na.rm = TRUE), 2), " — ", round(max(df[[var]], na.rm = TRUE), 2))
    } else if (is.factor(df[[var]])) {
      factor_levels = levels(df[[var]])
      paste(c(paste0(factor_levels[1],"*"), factor_levels[-1]), collapse = ", ")
    } else {
      paste(unique(df[[var]]), collapse = ", ")
    }
  }),
  N = sapply(rq5y, function(var) {
    sum(!is.na(df[[var]]))
  }),
  `Study Wave` = sapply(rq5y, get_study_wave)
)

# Combine all v_rq dataframes with row indices for grouping
v_rq1x_indexed = cbind(row_group = "RQ1 Variables", row_id = 1:nrow(v_rq1x), v_rq1x)
v_rq1y_indexed = cbind(row_group = "RQ1 Outcome Variables", row_id = (nrow(v_rq1x)+1):(nrow(v_rq1x)+nrow(v_rq1y)), v_rq1y)
v_rq2_indexed = cbind(row_group = "RQ2 Variables", row_id = (nrow(v_rq1x)+nrow(v_rq1y)+1):(nrow(v_rq1x)+nrow(v_rq1y)+nrow(v_rq2)), v_rq2)
v_rq3_indexed = cbind(row_group = "RQ3 Variables", row_id = (nrow(v_rq1x)+nrow(v_rq1y)+nrow(v_rq2)+1):(nrow(v_rq1x)+nrow(v_rq1y)+nrow(v_rq2)+nrow(v_rq3)), v_rq3)
v_rq5_indexed = cbind(row_group = "RQ5 Variables", row_id = (nrow(v_rq1x)+nrow(v_rq1y)+nrow(v_rq2)+nrow(v_rq3)+1):(nrow(v_rq1x)+nrow(v_rq1y)+nrow(v_rq2)+nrow(v_rq3)+nrow(v_rq5)), v_rq5)

v_rq_combined = rbind(v_rq1x_indexed, v_rq1y_indexed, v_rq2_indexed, v_rq3_indexed, v_rq5_indexed)

gt(v_rq_combined) %>%
  tab_row_group(
    label = "E) RQ4 Outcome Variables",
    rows = row_group == "RQ5 Variables"
  ) %>%
  tab_row_group(
    label = "D) RQ3 Outcome Variables",
    rows = row_group == "RQ3 Variables"
  ) %>%
  tab_row_group(
    label = "C) RQ2 Outcome Variables",
    rows = row_group == "RQ2 Variables"
  ) %>%
  tab_row_group(
    label = "B) RQ1-2 Participation Indicators (1 = yes, 0 = no)",
    rows = row_group == "RQ1 Outcome Variables"
  ) %>%
  tab_row_group(
    label = "A) Baseline variables used to predict later participation, and for IP weighting",
    rows = row_group == "RQ1 Variables"
  ) %>%
  cols_hide(c(row_group, row_id)) %>%
  cols_width(
    `Range.or.Level` ~ px(100)
  ) %>%
  cols_label_with(fn = ~ gsub("\\.", " ", .x)) %>%
  tab_style(
    style = cell_text(font = "Times New Roman", size = px(12)),
    locations = cells_body()
  ) %>%
  tab_style(
    style = cell_text(font = "Times New Roman", size = px(12)),
    locations = cells_column_labels()
  ) %>%
  tab_style(
    style = cell_text(font = "Times New Roman", size = px(12)),
    locations = cells_row_groups()
  ) %>%
  tab_footnote(
    footnote = "Note. MFQ = Mood and Feelings Questionnaire; SDQ = Strengths and Difficulties Questionnaire; GCSE = General Certificate of Secondary Education; KS3 = Key Stage 3; GAD-D = Generalized Anxiety Disorder scale; PARCA = Parent Report of Children's Abilities; MZ = Monozygotic; DZ = Dizygotic; HNC = Higher National Certificate; HND = Higher National Diploma; CSE = Certificate of Secondary Education."
  )

```

# Participation-Related plots

## Missing Data at each timepoint 

### Family-Level Participation Metrics

```{r}

df %>%
  select(c(acontact,rq1y)) %>%
  sapply(., function(x) length(which(x==1))/length(which(x>=0))) %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  `colnames<-`(c("Time Point","Participation Rate")) %>%
  mutate(`Time Point` = factor(`Time Point`, levels = `Time Point`, labels = var_to_label(`Time Point`))) %>%
  ggplot(aes(x = `Participation Rate`, y = `Time Point`)) + 
  geom_col() +
    geom_text(
    aes(label = paste0(round(`Participation Rate` * 100, 1), "%")),
    hjust = 1.1,  # Right align inside the bar (>1 moves it inside)
    color = "white",
    size = 3.5,
    fontface = "bold"
  ) +
  # geom_vline(aes(xintercept=.2)) +
  theme_bw()

save_plot("11_participation_rate", width = 9, height = 5)


df %>%
  select(cohort, acontact, all_of(rq1y)) %>%
  group_by(cohort) %>%
  summarise(across(everything(), 
                   ~ length(which(. == 1)) / length(which(. >= 0)),
                   .names = "{.col}"),
            .groups = "keep") %>%
  pivot_longer(cols = -cohort, names_to = "Time Point", values_to = "Participation Rate") %>%
  mutate(
    `Time Point` = factor(`Time Point`, 
                              levels = c("acontact", rq1y), 
                              labels = var_to_label(c("acontact", rq1y))),
         Time_numeric = as.numeric(`Time Point`),
    cohort = gsub(" twins", "",cohort),
    cohort = gsub(" to ", "-",cohort)
    ) %>%
  ggplot(aes(x = Time_numeric, y = `Participation Rate`, color = cohort, group = cohort)) + 
  geom_point(size = 3) +
  geom_line(linewidth = 1) +
  geom_text(
    aes(label = paste0(round(`Participation Rate` * 100, 1), "%")),
    vjust = -0.5,
    hjust = 0.5,
    size = 3,
    show.legend = FALSE
  ) +
  scale_x_continuous(
    breaks = 1:length(c("acontact", rq1y)),
    labels = var_to_label(c("acontact", rq1y))
  ) +
  scale_y_continuous(labels = scales::percent_format()) +
  theme_bw() +
  labs(x = "Time Point", 
       y = "Participation Rate", 
       color = "Cohort",
       title = "Participation Rates Over Time by Cohort") +
  theme(legend.position = c(.8,.8),
        axis.text.x = element_text(angle = 45, hjust = 1)) + 
  theme(plot.margin = margin(t = 1, r = 1, b = 1, l = 10, unit = "mm"))

save_plot("11_participation_rate_by_cohort", width = 11, height = 9)


df %>%
  select(cohort, acontact, all_of(rq1y)) %>%
  filter(btwoyear==1) %>%
  group_by(cohort) %>%
  summarise(across(everything(), 
                   ~ length(which(. == 1)) / length(which(. >= 0)),
                   .names = "{.col}"),
            .groups = "keep") %>%
  pivot_longer(cols = -cohort, names_to = "Time Point", values_to = "Participation Rate") %>%
  mutate(
    `Time Point` = factor(`Time Point`, 
                              levels = c("acontact", rq1y), 
                              labels = var_to_label(c("acontact", rq1y))),
         Time_numeric = as.numeric(`Time Point`),
    cohort = gsub(" twins", "",cohort),
    cohort = gsub(" to ", "-",cohort)
    ) %>%
  ggplot(aes(x = Time_numeric, y = `Participation Rate`, color = cohort, group = cohort)) + 
  geom_point(size = 3) +
  geom_line(linewidth = 1) +
  geom_text(
    aes(label = paste0(round(`Participation Rate` * 100, 1), "%")),
    vjust = -0.5,
    hjust = 0.5,
    size = 3,
    show.legend = FALSE
  ) +
  scale_x_continuous(
    breaks = 1:length(c("acontact", rq1y)),
    labels = var_to_label(c("acontact", rq1y))
  ) +
  scale_y_continuous(labels = scales::percent_format()) +
  theme_bw() +
  labs(x = "Time Point", 
       y = "Participation Rate", 
       color = "Cohort",
       title = "Participation Rates Over Time by Cohort",
       subtitle = "Just selecting participants that participated at YEAR 2"
       ) +
  theme(legend.position = c(.8,.8),
        axis.text.x = element_text(angle = 45, hjust = 1)) + 
  theme(plot.margin = margin(t = 1, r = 1, b = 1, l = 10, unit = "mm"))

save_plot("11_participation_rate_by_cohort_selectedy2", width = 11, height = 9)



```

### Twin-Level Participation Metrics

#### Participation Rates

```{r}

df %>%
  select(cohort, all_of(rq1y_twin1)) %>%
  # select(cohort, !starts_with(vars_to_na)) %>%
  group_by(cohort) %>%
  summarise(across(everything(), 
                   ~ length(which(. == 1)) / length(which(. >= 0)),
                   .names = "{.col}"),
            .groups = "keep") %>%
  pivot_longer(cols = -cohort, names_to = "Time Point", values_to = "Participation Rate") %>%
  mutate(
    `Time Point` = factor(`Time Point`, 
                              levels = paste0(rq1y_twin,"1"), 
                              labels = rq1y_twin_labels),
    # Time_numeric = as.numeric(`Time Point`),
    cohort = gsub(" twins", "",cohort),
    cohort = gsub(" to ", "-",cohort)
    ) %>%
  ggplot(aes(x = `Time Point`, y = `Participation Rate`, color = cohort, group = cohort)) + 
  geom_point(size = 3) +
  geom_line(linewidth = 1) +
  geom_text(
    aes(label = paste0(round(`Participation Rate` * 100, 1), "%")),
    vjust = -0.5,
    hjust = 0.5,
    size = 3,
    show.legend = FALSE
  ) +
  # scale_x_continuous(
  #   breaks = 1:length(c("acontact", rq1y)),
  #   labels = var_to_label(c("acontact", rq1y))
  # ) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0,1)) +
  theme_bw() +
  labs(x = "Time Point", 
       y = "Participation Rate", 
       color = "Cohort",
       title = "Participation Rates Over Time by Cohort") +
  theme(legend.position = c(.8,.8),
        axis.text.x = element_text(angle = 45, hjust = 1)) + 
  theme(plot.margin = margin(t = 1, r = 1, b = 1, l = 10, unit = "mm")) 

```


#### Twin Concordance

NOTE THAT WE REMOVE COHORTS 2 & 3

ALSO FOR CORRELATIONS - I SHOULD FILTER BY TWIN == 1! 

```{r}

vars_to_na = c("btwdata","ctwdata", "gcdata","icdata","jcdata","pcwebdata")

df_cohort34_removed = df %>%
  mutate(
      across(starts_with(vars_to_na), ~ ifelse(cohort %in% c("Cohort 3: twins born Sep-95 to Aug-96","Cohort 4: twins born Sep-96 to Dec-96"), NA, .))
    )

plot_data = map_dfr(rq1y_twin, function(var){
  df_cohort34_removed %>% 
    group_by(sexzyg) %>%
    summarise(
      mean = mean(.data[[paste0(var,"1")]], na.rm = TRUE),
      cor  = cor(x=.data[[paste0(var,"1")]], y=.data[[paste0(var,"2")]], use = "pairwise.complete.obs")
    # cor  = {
    #   group_data <- cur_data()
    #   as.numeric(psych::tetrachoric(x=group_data[paste0(var,1:2)])$rho)[2]
    # }
    ) %>%
    ungroup() %>%
    mutate(variable = var)
}, .id = "var_index") %>%
  pivot_longer(cols = c(mean, cor), 
               names_to = "metric", 
               values_to = "value") %>%
  mutate(
    metric_label = case_when(
      metric == "mean" ~ "Participation Rate",
      metric == "cor" ~ "Twin Tetrachoric Correlation"
    ),
    time_point = rq1y_twin_labels_clean[match(.$variable, rq1y_twin)],
    time_point = factor(time_point, levels = rq1y_twin_labels_clean)
  )

plot_data %>%
  ggplot( aes(x = time_point, y = value, color = sexzyg, group = sexzyg)) +
  geom_line(size = 1, alpha = 0.8) +
  geom_point(size = 2.5, alpha = 0.9) +
  ggrepel::geom_text_repel(aes(label = case_when(
    metric == "mean" ~ paste0(round(value * 100, 1), "%"),
    metric == "cor" ~ gbtoolbox::apa_num(value)
  )),
  position = position_dodge(width = 0.5),
  vjust = 2,
  size = 2,
  show.legend = FALSE) +
  facet_wrap(~metric_label, scales = "free_y", ncol = 1) +
  scale_color_manual(
    values = c(
      "MZ male" = "#d7191c",
      "DZ male" = "#fdae61",
      "MZ female" = "#2c7bb6",
      "DZ female" = "#abd9e9",
      "DZ opposite sexes" = "grey"
    )
  ) +
  ggh4x::scale_y_facet(metric_label == "Participation Rate",
                       labels = scales::percent_format(),
                       breaks = seq(0, 1, 0.1)) +
  labs(
    # title = "Twin Study: Participation Rates and Correlations Across Time Points",
    # subtitle = "Trends by Zygosity and Sex",
    x = "Participation Time Point",
    y = NULL,
    color = "Zygosity",
  ) +
  theme_minimal() +
  theme(
    # plot.title = element_text(size = 14, face = "bold"),
    # plot.subtitle = element_text(size = 12),
    # strip.text = element_text(size = 11, face = "bold"),
    legend.spacing.y = unit(0, "pt"),
    legend.background = element_rect(fill = alpha("white",.5), color = NA),
    legend.key.height = unit(0.6, "pt"),
    legend.text = element_text(margin = margin(t = 0, b = 0)),
    legend.position = c(0.02, 0.02),
    legend.justification = c(0, 0),
    # legend.title = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 47, unit = "pt")
  ) +
  guides(color = guide_legend(override.aes = list(size = 3)))

save_plot("11_participation_rates_zygosity", width = 7, height = 6)


```

#### Create ACE estimates plot 

```{r}
estimates = readRDS(file.path("results","1_ace_estimates.Rds"))
ci       = readRDS(file.path("results","1_ace_confints.Rds"))

estimates %>%
    mutate(
      par_clean = case_when(
        par == "a1" ~ "A",
        par == "c1" ~ "C",
        par == "e1" ~ "E"
      ),
      time_point = rq1y_twin_labels_clean[match(name, rq1y_twin)],
      time_point = factor(time_point, levels = rq1y_twin_labels_clean)
    ) %>%
    ggplot(aes(x = time_point, y = value, color = par_clean, group = interaction(sex, par_clean))) +
    geom_line(size = .5, alpha = 0.8, aes(linetype = sex)) +
    geom_point(size = 2.5, alpha = 0.9, aes(shape = sex)) +
    # ggrepel::geom_text_repel(aes(label = sprintf("%.2f", value)),
    #   size = 1.5,
    #   show.legend = FALSE,
    #   min.segment.length = 0,
    #   segment.size = 0.2,
    #   max.overlaps = Inf,
    #   direction = "y",
    #   force = 2) +
    scale_color_manual(
      values = c(
        "A" = "#d7191c",
        "C" = "#fdae61", 
        "E" = "#2c7bb6"
      )
    ) +
    scale_linetype_manual(
      values = c("male" = "solid", "female" = "dotted")
    ) +
    scale_shape_manual(
      values = c("male" = 16, "female" = 17)
    ) +
    scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
    labs(
      x = "Participation Time Point",
      y = NULL,
      color = "ACE Component",
      linetype = "Sex",
      shape = "Sex"
    ) +
    theme_minimal() +
    theme(
      legend.background = element_blank(),
      legend.position = "top",
      legend.justification = "center",
      legend.direction = "horizontal",
      legend.box = "horizontal",
      legend.key.height = unit(0.5, "lines"),
      legend.key.width = unit(0.8, "lines"),
      legend.text = element_text(size = 8),
      legend.title = element_text(size = 9),
      legend.margin = margin(2, 2, 2, 2),
      panel.grid.minor = element_blank(),
      axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
      plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = 47, unit = "pt"),
      plot.title = element_text(size = 8, hjust = 0.5),
      strip.text = element_text(size = 10, face = "bold")
    ) +
    guides(
      color = guide_legend(override.aes = list(size = 2), order = 1),
      linetype = guide_legend(override.aes = list(size = 0.8), order = 2),
      shape = guide_legend(override.aes = list(size = 1.5), order = 2)
    )

  save_plot("11_ACE_time", width = 7, height = 3)



```


#### ANOVA Sex & Zygosity predicting participation

```{r}

map_dfr(seq_along(rq1y_twin1), function(i){
  y_var     = rq1y_twin1[i]
  model     = glm(paste(y_var, "~ zygos + sex1"), data = df, family = binomial)
  robust_se = clubSandwich::coef_test(model, cluster = df$randomfamid, vcov = "CR0", test = "z")
  
  data.frame(
    `Time Point`    = rq1y_twin_labels_clean[i],
    `Zygosity Coef` = as.numeric(robust_se[2, "beta"]),
    `Zygosity p`    = as.numeric(robust_se[2, "p_z"]),
    `Sex Coef`      = as.numeric(robust_se[3, "beta"]),
    `Sex p`         = as.numeric(robust_se[3, "p_z"])
  )
}) %>%
  knitr::kable(digits = 3)

```


## How many different time points do we have for each participant (excluding first contact)

```{r}

df %>% 
  select(c(rq1y)) %>% 
  apply(., 1, function(x) sum(x)) %>% 
  tibble(`Number Time Points Participated` = .) %>% 
  count(`Number Time Points Participated`) %>% 
  ggplot(aes(x = `Number Time Points Participated`, y = n)) + 
  geom_col() + 
  geom_text(aes(label = n), vjust = 1.2, color = "white", size = 3.5) +
  theme_bw() + 
  labs(
    title = "How many times has each pps been tested (exl. first contact)?",
    subtitle = "Twin-level participation numbers (N = 27890)"
    )

save_plot("11_number_of_times_each_twin_participated", width = 9, height = 5)

```
## How predictive is cohort and year of birth in participation at each timepoint? 

I've checked there are no missing values for cohort or year of birth (aonsby)

[View full-sizeplot](plots/11_r2_participation_cohort_birthyear.pdf){target="_blank"}

```{r}
models = list()

for (i in seq_along(rq1y)){
  # cat(i, "/", length(rq1y),"\n")
  formula <- as.formula(paste(rq1y[i], "~ cohort + aonsby"))
  models[[i]] = glm(formula, data = df, family = binomial)
}

var_explained = sapply(models, function(x) as.numeric(performance::r2(x)))

cbind.data.frame(rq1y, rq1y_labels, var_explained) %>%
  `colnames<-`(c("Var","Label","R2 by Cohort and Birth Year")) %>%
  knitr::kable(digits = 2)


cbind.data.frame(rq1y, rq1y_labels, var_explained) %>%
  `colnames<-`(c("Var","Label","Variance Explained")) %>%
  mutate(
    Label = factor(Label, levels = rq1y_labels)
  ) %>%
  ggplot(aes(x=Label,y=`Variance Explained`)) + 
  geom_col() + 
  geom_text(aes(label = gbtoolbox::apa_num(`Variance Explained`)),
            vjust = 1.1, color = "white", size = 3) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1, vjust=1)) +
  scale_x_discrete(expand = expansion(mult = c(0.1, 0)))

save_plot("11_r2_participation_cohort_birthyear", width = 9, height = 5)


```

At the Y18 time point, we see that those in cohort 1 and those born in 1994 are less likely to participate. 

```{r}

rq1y[10]

var_to_label(rq1y[10])

summary(models[[10]])

```



### Cohort alone is not helpful for determining who was eligible at t2 and t3

As we can see below, nobody in cohort 4 took part in Y2, Y3, Y10 data collection. However, a minority of participants took part in Y2-Y3 from cohort 3, but participation rates are quite low suggesting that some participants were elibile and some not. 

```{r}

 table(df0$cohort, df0$btwoyear)
 table(df0$cohort, df0$cthreeyr)
 table(df0$cohort, df0$jtenyear)

```

```{r}

 table(df0$AONSBY, df0$btwoyear)
 table(df0$AONSBY, df0$cthreeyr)
 table(df0$AONSBY, df0$jtenyear)

```




## Participation Patterns

```{r}

df %>%
  select(c(rq1y)) %>% 
  rename_with(~var_to_label(., df)) %>%
  mutate_all(~na_if(., 0)) %>%
  ggmice::plot_pattern(npat = 70, rotate = TRUE) + 
  scale_x_discrete(limits = (var_to_label(rq1y))) + 
  coord_flip()

save_plot("11_participation_pattern", width = 19, height = 6)


df %>%
  select(c(rq1y)) %>%
  mutate_all(~ .==1) %>%
  rowwise() %>%
  mutate(
    `Year 2 - 4 present`   = any(btwoyear, cthreeyr, dfouryr),
    `Year 7 - 12 present`  = any(gsevenyr, heightyr, jtenyear),
    `Year 14 - 16 present` = any(n14year, p16year),
    `Year 18 - 26 present` = any(rcqdata, uteds21data, zmhdata)
  ) %>%
  select(ends_with("present")) %>%
  mutate_all(~ na_if(., FALSE)) %>%
  ggmice::plot_pattern(rotate = TRUE) + 
  scale_x_discrete(limits = c("Year 2 - 4 present", "Year 7 - 12 present", "Year 14 - 16 present", "Year 18 - 26 present")) + 
  coord_flip()

save_plot("11_participation_pattern_grouped", width = 19, height = 6)

df %>%
  select(c(rq1y)) %>%
  rename_with(~var_to_label(., df)) %>%
  rename_with(~clean_rq1y_label(.)) %>%
  mutate_all(~ na_if(., 0)) %>%
  gbtoolbox::plot_pairwise_missing(textadjust = 2.5) +
  labs(subtitle = "Number of twins tested at each time point (diagonal) and \npair of time points (below diagonal)")

save_plot("11_pairwise_missingness_plot", width = 6, height = 6)

```

# Descriptive stats on specific variables

## Age of twins at each time point

[View full-size plot](plots/11_participant_ages_each_timepoint.pdf){target="_blank"}


```{r}

df_age = df %>%
  select(contains("age")) %>%
  select(-contains("genpro"), -contains("mumage"), - contains("dadage"), -ends_with("2")) 

df_age_long = df_age %>%
  `colnames<-`(var_to_label(colnames(df_age))) %>%
  pivot_longer(cols = everything())

df_age_long %>% 
  mutate(name = factor(name, levels = var_to_label(colnames(df_age)))) %>%
  ggplot(aes(x=value)) + 
  geom_histogram(bins = 100) +
  facet_wrap(~name, ncol =3) + 
  scale_x_continuous(breaks = seq(0,30,by=2)) +
  theme_bw() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
    )

save_plot("11_participant_ages_each_timepoint", width = 12, height = 12)


```

## Education

```{r}

all(table(df$amohqual)==table(df$amohqualn))

df %>%
  filter(twin ==1) %>%
  count(amohqual) %>%
  mutate(
    percent = n/sum(n)
  ) %>%
  gt() %>% 
  fmt_percent(columns = percent)


```

## Relationship between maternal education and child outcomes

This section examines whether maternal education level predicts child academic and cognitive outcomes. We regress each outcome on maternal education using dummy coding (treating each education level as a separate category), which allows us to detect non-linear relationships.

```{r}

# Helper function to create marginal effects plot
plot_education_effects = function(model, y_label) {
  marginaleffects::predictions(model, by = "amohqual") %>%
    data.frame() %>%
    rename(
      prediction = estimate,
      education_level = amohqual
    ) %>%
    ggplot(aes(y = prediction, x = education_level, group = 1)) +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
    geom_point(size = 2) +
    geom_line() +
    labs(
      x = "Maternal Education Level",
      y = y_label
    ) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
    )
}

```

### KS3 Academic Achievement (Age 14)

```{r}

model_ks3 = df %>%
  filter(twin == 1) %>%
  filter(!is.na(amohqual)) %>%
  lm(npks3tall1 ~ amohqual, data = .)

summary(model_ks3)

plot_education_effects(model_ks3, "Predicted KS3 Score")

save_plot("11_meducation_ks3_plot", width = 6, height = 5)

```

### GCSE Core Subjects Grade (Age 16)

```{r}

model_gcse = df %>%
  filter(twin == 1) %>%
  filter(!is.na(amohqual)) %>%
  lm(pcexgcsecoregrdm1 ~ amohqual, data = .)

summary(model_gcse)

plot_education_effects(model_gcse, "Predicted GCSE Score")

save_plot("11_meducation_gcse_plot", width = 6, height = 5)

```

### Cognitive Ability (Age 12)

```{r}

model_cog = df %>%
  filter(twin == 1) %>%
  filter(!is.na(amohqual)) %>%
  lm(lcg1 ~ amohqual, data = .)

summary(model_cog)

plot_education_effects(model_cog, "Predicted Cognitive Score (g)")

save_plot("11_meducation_cognitive_plot", width = 6, height = 5)

```