---
title: "RQ2: Artificially Attritioned Datasets"
format: 
  html:
    embed-resources: true
    fig-width: 8
    fig-height: 8
    code-link: true
    code-fold: true
    code-tools: true
    df-print: paged
    toc: true
    grid:
      body-width: 900px
editor: source
editor_options: 
  chunk_output_type: console
---

# Data cleaning

## Load Results

The script [2_rq2_1_run_bootstrapping.R](2_rq2_1_run_bootstrapping.R){target="_blank"} runs the bootstrap analyses and saves the results - which are then cleaned and presented below. 

```{r}
#| output: false

source("0_load_data.R")
rm(rq1y, rq1y_labels, rq1y_labels_clean)

# Load new bootstrap results
boot_results           = readRDS(file = file.path("results", "2_boot_results.Rds"))

number_timepoints = length(rq1y_twin1)
number_bootstraps = length(boot_results)/number_timepoints

# ||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||
# Post-processing of results ---------------------------------------------------
# ||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||

joblist = expand.grid(
  b = 1:number_bootstraps,
  i = 1:number_timepoints
  )

### Extract data from boot_results ---------------------------------------------

md_df = do.call(rbind, lapply(seq_along(boot_results), function(j) {
  result_vec = boot_results[[j]]$md[[1]]
  data.frame(
    .dataset = joblist[j, "i"],
    .boot    = joblist[j, "b"],
    t(setNames(result_vec, rq2y))
  )
}))

smd_df = do.call(rbind, lapply(seq_along(boot_results), function(j) {
  result_vec = boot_results[[j]]$smd[[1]]
  data.frame(
    .dataset = joblist[j, "i"],
    .boot    = joblist[j, "b"],
    t(setNames(result_vec, rq2y))
  )
}))

var_df = do.call(rbind, lapply(seq_along(boot_results), function(j) {
  result_vec = boot_results[[j]]$var[[1]]
  data.frame(
    .dataset = joblist[j, "i"],
    .boot    = joblist[j, "b"],
    t(setNames(result_vec, rq2y))
  )
}))

cor_df = do.call(rbind, lapply(seq_along(boot_results), function(j) {
  result_vec = boot_results[[j]]$cor_resid[[1]]
  data.frame(
    .dataset = joblist[j, "i"],
    .boot    = joblist[j, "b"],
    t(result_vec)
  )
}))

srmr_df = do.call(rbind, lapply(seq_along(boot_results), function(j) {
  data.frame(
    .dataset = joblist[j, "i"],
    .boot    = joblist[j, "b"],
    srmr     = boot_results[[j]]$srmr[[1]]
  )
}))

ace_df = do.call(rbind, lapply(seq_along(boot_results), function(j) {
  ace_est  = boot_results[[j]]$ace[[1]]$ace_estimates
  ace_diff = boot_results[[j]]$ace[[1]]$ace_differences
  result_df = rbind(ace_est, ace_diff)
  result_df$.dataset = joblist[j, "i"]
  result_df$.boot    = joblist[j, "b"]
  return(result_df)
}))

rownames(md_df)   = NULL
rownames(smd_df)  = NULL
rownames(var_df)  = NULL
rownames(cor_df)  = NULL
rownames(srmr_df) = NULL

bootstrap_iter = list(md = md_df, smd = smd_df, var = var_df, cor_resid = cor_df, srmr = srmr_df)

### Calculate p-values and confidence intervals per dataset --------------------

bootstrap_summary_df = lapply(names(bootstrap_iter)[1:3], function(stat_name) {
  df = bootstrap_iter[[stat_name]]
  
  dataset = df %>%
    pivot_longer(
      cols = !starts_with(".")
    ) %>%
    group_by(.dataset, name) %>%
    summarise(
      .mean_qi_pd(value),
      .groups = "drop"
    ) %>%
    mutate(
      dataset       = rq1y_twin_labels_clean_extrashort[.dataset],
      variable      = rq2y_labels_short[match(name, rq2y)],
      stat          = stat_name
    ) %>%
    select(-.dataset, -name) %>%                                                                                                                                                           
    relocate(dataset, variable, stat) 
})

bootstrap_summary_df = do.call(rbind, bootstrap_summary_df)
rownames(bootstrap_summary_df) = NULL

### Calculate bootstrap summaries for ACE separately ---------------------------

bootstrap_summary_df_ace = ace_df %>%
  filter(name != "amohqualn1") %>%
  mutate(
    group = recode(group,
      "df1"  = "Attritioned",
      "df2"  = "Original",
      "diff" = "Difference"
    ),
    dataset       = rq1y_twin1[.dataset],
    dataset_label = rq1y_twin_labels_clean_extrashort[.dataset]
  ) %>%
  group_by(par, name, sex, group, dataset, dataset_label) %>%
  summarise(
    .mean_qi_pd(value),
    .groups = "drop"
  ) 


variable_comparisons_df = bootstrap_summary_df %>%
  group_by(dataset, stat) %>%
  mutate(
    pval_adj = stats::p.adjust(pval, method = "holm")
  )

```

## Create Attritioned Datasets

To remove participants, I've set the rows to NA rather than remove the rows, so all the datasets have the same length. This preserves the paired twin structure needed for bootstrapping and ACE analyses.

```{r}

attritioned_datasets = list()

for (i in seq_along(rq1y_twin1)){
  
  filter = as.numeric(df[[rq1y_twin1[i]]])==0 # 1 = present (Y), 2 = not-present (N)
  
  attritioned_datasets[[i]] = df %>% 
    select(all_of(rq2y))
  
  attritioned_datasets[[i]][filter,] = NA
  
}

original_dataset = df %>%
  select(all_of(rq2y))


```

# Comparison of Means, Standardized Mean Differences, and Variances 

## Create GT Results table

```{r}

variable_comparisons_df %>% 
  select(-starts_with("."),-pd, -pval) %>%
  pivot_wider(
    names_from = "stat",
    values_from = c("y","ymin","ymax","pval_adj"),
    id_cols = c("variable", "dataset")
  ) %>%
  dplyr::rename( Outcome = "variable") %>%
  gt(groupname_col = "dataset") %>%
  tab_options(
    table.width = px(800)
    # table.border.top.style = "none",
    # table.border.bottom.style = "none",
    # column_labels.border.bottom.style = "solid",
    # column_labels.border.bottom.width = px(1),
    # column_labels.border.bottom.color = "black",
    # table_body.border.bottom.style = "none",
    # table_body.vlines.style = "none",
    # row_group.border.bottom.style = "none"
  ) %>%
  
  # Create column spanners (nested headers)
  tab_spanner(
    label = "Mean Difference",
    columns = c(y_md, ymin_md, ymax_md, pval_adj_md)
  ) %>%
  tab_spanner(
    label = "Standardized Mean Difference", 
    columns = c(y_smd, ymin_smd, ymax_smd, pval_adj_smd)
  ) %>%
  tab_spanner(
    label = "Variance % Change",
    columns = c(y_var, ymin_var, ymax_var, pval_adj_var)
  ) %>%
  
  # Rename columns under each spanner
  cols_label(
    y_md = "Est", ymin_md = "LB", ymax_md = "UB", pval_adj_md = "p",
    y_smd = "Est", ymin_smd = "LB", ymax_smd = "UB", pval_adj_smd = "p", 
    y_var = "Est", ymin_var = "LB", ymax_var = "UB", pval_adj_var = "p"
  ) %>%
  
  # Format numbers
  fmt(
    columns = !contains("var") & !contains("Outcome"),
    fns = function(x) {gbtoolbox::apa_num(x, n_decimal_places = 3)}
  ) %>% 
  fmt(
    columns = "pval_adj_var",
    fns = function(x) {gbtoolbox::apa_num(x, n_decimal_places = 3)}
  ) %>% 
  fmt_percent(
    columns = c("y_var", "ymin_var","ymax_var"),
    decimals = 2,
    drop_trailing_zeros = FALSE,
    drop_trailing_dec_mark = FALSE
  ) %>%
  # Styling - uniform font size
  tab_style(
    style = cell_text(size = px(10)),
    locations = cells_column_spanners()
  ) %>%
  tab_style(
    style = cell_text(size = px(10)),
    locations = cells_body()
  ) %>%
  tab_style(
    style = cell_text(size = px(10), align = "center"),
    locations = cells_row_groups()
  ) %>%
  tab_style(
    style = cell_text(size = px(10)),
    locations = cells_column_labels()
  ) %>%
  tab_style(
    style = cell_text(size = px(10)),
    locations = cells_footnotes()
  ) %>%
  
  # Highlight significant results - positive effects (light green)
  tab_style(
    style = cell_fill(color = "#d5e8d4"),  # Light green for positive effects
    locations = cells_body(
      columns = c(y_md, ymin_md, ymax_md, pval_adj_md),
      rows = pval_adj_md < 0.05 & y_md > 0
    )
  ) %>%
  tab_style(
    style = cell_fill(color = "#d5e8d4"),  # Light green for positive effects
    locations = cells_body(
      columns = c(y_smd, ymin_smd, ymax_smd, pval_adj_smd),
      rows = pval_adj_smd < 0.05 & y_smd > 0
    )
  ) %>%
  tab_style(
    style = cell_fill(color = "#d5e8d4"),  # Light green for positive effects
    locations = cells_body(
      columns = c(y_var, ymin_var, ymax_var, pval_adj_var),
      rows = pval_adj_var < 0.05 & y_var > 0
    )
  ) %>%
  
  # Highlight significant results - negative effects (light red)
  tab_style(
    style = cell_fill(color = "#f8cecc"),  # Light red for negative effects
    locations = cells_body(
      columns = c(y_md, ymin_md, ymax_md, pval_adj_md),
      rows = pval_adj_md < 0.05 & y_md < 0
    )
  ) %>%
  tab_style(
    style = cell_fill(color = "#f8cecc"),  # Light red for negative effects
    locations = cells_body(
      columns = c(y_smd, ymin_smd, ymax_smd, pval_adj_smd),
      rows = pval_adj_smd < 0.05 & y_smd < 0
    )
  ) %>%
  tab_style(
    style = cell_fill(color = "#f8cecc"),  # Light red for negative effects
    locations = cells_body(
      columns = c(y_var, ymin_var, ymax_var, pval_adj_var),
      rows = pval_adj_var < 0.05 & y_var < 0
    )
  ) %>%
  
  # Add footnotes
  tab_footnote(
    footnote = md("<em>Note.</em> Est = Estimate, LB = Lower Bound 95% Confidence Interval, UB = Upper Bound 95% Confidence Interval. Significant (p<sub>Bonferroni-Holm</sub>) effects are highlighted in green (increases) or red (decreases)."),
    # locations = cells_column_spanners(),
    placement = "right"
  ) %>%
  tab_footnote(
    footnote = "P values are Bonferroni-Holm adjusted within each participation group",
    locations = cells_column_labels(columns = contains("pval_adj")),
    placement = "right"
  ) %>%
  tab_footnote(
    footnote = md("Standardized mean difference = (Mean<sub>original</sub> - Mean<sub>attritioned</sub>) / SD<sub>original</sub>"),
    locations = cells_column_spanners(spanners = "Standardized Mean Difference"),
    placement = "right"
  ) %>%
    tab_footnote(
    footnote = md("Percentage change = (Var<sub>attritioned</sub> - Var<sub>original</sub>) / Var<sub>original</sub>"),
    locations = cells_column_spanners(spanners = "Variance % Change"),
    placement = "right"
  ) %>%
  opt_footnote_marks(marks = c("*", "†","‡"))


```

## Create circular heatmaps for changes in means and variances

### Standardised Mean Differences 

```{r}
  # Create circular heatmap for SMD
   heatmap_data_smd = variable_comparisons_df %>%
    filter(stat == "smd") %>%
    mutate(
      outcome_labeled = dataset,
      outcome_labeled = factor(outcome_labeled, levels = rq1y_twin_labels_clean_extrashort),
      Variables_wrapped = str_wrap(variable, width = 8),
      # Set fill value to NA if not significant, otherwise keep original sign
      fill_value = case_when(
        pval_adj < 0.05 ~ y,
        TRUE ~ NA
      )
    ) %>%
    # Order variables by number of significant effects
    group_by(variable) %>%
    mutate(n_significant = sum(pval_adj < 0.05)) %>%
    ungroup() %>%
    mutate(
      Variables_wrapped = factor(Variables_wrapped,
                                levels = unique(Variables_wrapped[order(-n_significant)]))
    )

  ggplot(heatmap_data_smd, aes(x = Variables_wrapped, y = outcome_labeled)) +
    geom_tile(aes(fill = fill_value, 
                  color = ifelse(variable %in% c("Emotional problems", "Prosocial behavior"), "azure2", "black")), 
              size = 0.5) +
    scale_color_identity() +
    geom_text(
      aes(label = ifelse(pval_adj < 0.05,
                         gsub("^(-?)0\\.", "\\1.", sprintf("%.3f", y)), ""),
          size = ifelse(outcome_labeled == "Y4", 2, 3.5)
      ),
      color = "black"
    ) +
    scale_size_identity() +
    coord_polar(theta = "x", start = 0, direction = 1, clip = "off") +
    scale_fill_gradient2(
      low = "#2166ac", 
      mid = "white", 
      high = "#d73027",
      midpoint = 0,
      # limits = c(-.1, 
      #            max(heatmap_data_smd$fill_value, na.rm = TRUE)),
      na.value = "white",
      guide = "none"
    ) +
    labs(
      title = "Standardised Mean Difference",
      subtitle = expression(frac(bar(X)[attritioned] - bar(X)[original], SD[original])),
      tag = "A"
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 0, size = 11, colour = "black"),
      axis.text.y = element_blank(),
      axis.title = element_blank(),
      panel.grid = element_blank(),
      panel.background = element_blank(),
      plot.background = element_blank(),
      plot.title = element_text(hjust = 0.5, size = 16),
      plot.subtitle = element_text(hjust = 0.5, size = 13.5, margin = margin(b = -20)),
      plot.tag = element_text(hjust = 0, vjust = 0, size = 30, face = "bold"),
      plot.tag.position = "topleft",
      plot.margin = margin(2, 10, -33, 10)
    ) +
    {
      y_positions = seq_along(levels(heatmap_data_smd$outcome_labeled))
      labels = levels(heatmap_data_smd$outcome_labeled)

      lapply(seq_along(labels), function(i) {
        annotate("text", x = 0.5, y = y_positions[i], 
                 label = paste0(labels[i], " "), 
                 size = 3.7, 
                 hjust = 1, angle = 7)
      })
    }

  save_plot("2_rq2_circular_heatmap_smd", width = 8, height = 8)





```

### Variance Differences

```{r}
  # Create circular heatmap for variance differences
   heatmap_data_var = variable_comparisons_df %>%
    filter(stat == "var") %>%
    mutate(
      outcome_labeled = dataset,
      outcome_labeled = factor(outcome_labeled, levels = rq1y_twin_labels_clean_extrashort),
      Variables_wrapped = str_wrap(variable, width = 8),
      # Set fill value to NA if not significant, otherwise keep original sign
      fill_value = case_when(
        pval_adj < 0.05 ~ y,
        TRUE ~ NA
      )
    ) %>%
    # Order variables by number of significant effects
    group_by(variable) %>%
    mutate(n_significant = sum(pval_adj < 0.05)) %>%
    ungroup() %>%
    mutate(
      Variables_wrapped = factor(Variables_wrapped,
                                levels = unique(Variables_wrapped[order(-n_significant)]))
    )

  ggplot(heatmap_data_var, aes(x = Variables_wrapped, y = outcome_labeled)) +
    geom_tile(aes(fill = fill_value, 
                  color = ifelse(variable %in% c("Emotional problems", "Prosocial behavior", "Hyperactivity", "Grammar"), "azure2", "black")), 
              size = 0.5) +
    scale_color_identity() +
    geom_text(
      aes(label = ifelse(pval_adj < 0.05,
                         paste0(sprintf("%.1f", y * 100), "%"), ""),
          size = ifelse(outcome_labeled == "Y4", 2, 3.5)
      ),
      color = "black"
    ) +
    scale_size_identity() +
    coord_polar(theta = "x", start = 0, direction = 1, clip = "off") +
    scale_fill_gradient2(
      low = "#2166ac", 
      mid = "white", 
      high = "#d73027",
      midpoint = 0,
      limits = c(-max(abs(heatmap_data_var$fill_value), na.rm = TRUE), 
                 max(abs(heatmap_data_var$fill_value), na.rm = TRUE)),
      na.value = "white",
      guide = "none"
    ) +
    labs(
      title = "Variance Percentage Change",
      subtitle = expression(frac(Var[attritioned] - Var[original], Var[original]) %*% 100),
      tag = "B"
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 0, size = 11, colour = "black"),
      axis.text.y = element_blank(),
      axis.title = element_blank(),
      panel.grid = element_blank(),
      panel.background = element_blank(),
      plot.background = element_blank(),
      plot.title = element_text(hjust = 0.5, size = 16),
      plot.subtitle = element_text(hjust = 0.5, size = 13.5, margin = margin(b = -20)),
      plot.tag = element_text(hjust = 0, vjust = 0, size = 30, face = "bold"),
      plot.tag.position = "topleft",
      plot.margin = margin(2, 10, -33, 10)
    ) +
    {
      y_positions = seq_along(levels(heatmap_data_var$outcome_labeled))
      labels = levels(heatmap_data_var$outcome_labeled)

      lapply(seq_along(labels), function(i) {
        annotate("text", x = 0.5, y = y_positions[i],
                 label = paste0(labels[i], " "), size = 3.7, hjust = 1, angle = 7)
      })
    }

  save_plot("2_rq2_circular_heatmap_var", width = 8, height = 8)

```


## Plot data distributions

```{r}

attritioned_datasets_long = attritioned_datasets

for (i in seq_along(attritioned_datasets_long)){
  attritioned_datasets_long[[i]] = gather(as.data.frame(as.matrix(attritioned_datasets[[i]])))
  attritioned_datasets_long[[i]]$dataset = rq1y_twin1[i]
}

x = original_dataset %>%
    as.matrix() %>% as.data.frame() %>%
    gather() %>%
    mutate(dataset = "original")

attritioned_datasets_long = c(attritioned_datasets_long, list(x))

attritioned_datasets_long = do.call(rbind.data.frame,attritioned_datasets_long)

attritioned_datasets_long = attritioned_datasets_long %>%
  filter(!is.na(value)) %>%
  mutate(
    key = rq2y_labels_short[match(key, rq2y)]
  )
  

attritioned_datasets_long %>%
  mutate(dataset = factor(dataset)) %>%
  ggplot(aes(x = value, group = dataset, col = dataset)) +
  geom_density() +
  facet_wrap(~key, scales = "free")

```

### Alternative: Frequency polygon (better for ordinal/integer data)

Only comparing a single attrition timepoint (Y26 Questionnaire). 

```{r}
#| fig-width: 12
#| fig-height: 10

attritioned_datasets_long %>%
  filter(!(key %in% c("Parent-admin cognition","Vocabulary"))) %>% 
  filter(dataset %in% c("original", "zmhdata1")) %>%
  mutate(
    dataset = factor(dataset),
    dataset = ifelse(dataset == "original", "original", "attritioned"),
    value = as.numeric(value)
  ) %>%
  count(key, value, dataset) %>%
  group_by(key, dataset) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = value, y = prop, col = dataset, group = dataset)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_color_manual(values = c("original" = "steelblue", "attritioned" = "coral")) +
  facet_wrap(~key, scales = "free") +
  labs(
    x = "Value",
    y = "Proportion",
    title = "Distribution comparison: Original vs Participants at Y26 Questionnaire"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

```

# Correlations

The following code creates `x_var` and `y_var` vectors that map correlation column indices (X1, X2, etc.) to variable pairs from the lower triangle of the correlation matrix.

```{r}
test_correlation_matrix = matrix(
  nrow = length(rq2y),
  ncol = length(rq2y)
  )

for(i in seq_along(rq2y)){
  for(j in seq_along(rq2y)){
    test_correlation_matrix[i,j] = paste(rq2y[i], rq2y[j], collapse = " ")
  }
}

x = test_correlation_matrix[lower.tri(test_correlation_matrix)]

x_var = str_extract(x, "^\\S+")
y_var = str_extract(x, "\\S+$")

rm(test_correlation_matrix,x)
```

## Plot changes in correlations

```{r}

# Get original column order before pivot
cor_col_order = setdiff(names(cor_df), c(".dataset", ".boot"))

correlation_comparisons_df = cor_df %>%
  pivot_longer(
    cols = !starts_with(".")
  ) %>%
  mutate(
    name = factor(name, levels = cor_col_order)
  ) %>%
  group_by(.dataset, name) %>%
  summarise(
    .mean_qi_pd(value),
    .groups = "drop"
  ) %>%
  group_by(.dataset) %>%
  mutate(
    pval_adj = stats::p.adjust(pval, method = "holm")
  ) %>%
  ungroup() %>%
  mutate(
    x_var = x_var[match(name, cor_col_order)],
    y_var = y_var[match(name, cor_col_order)]
  ) %>% 
  arrange(.dataset, name)

correlation_comparisons = split(correlation_comparisons_df, correlation_comparisons_df$.dataset)
correlation_comparisons = lapply(correlation_comparisons, function(x) select(x, -.dataset))
names(correlation_comparisons) = rq1y_twin1[as.numeric(names(correlation_comparisons))]

library(patchwork)

plots = lapply(1:length(correlation_comparisons), function(i) {
  correlation_comparisons[[i]] %>%
    plot_lower_triangular_matrix2(
      variables = rq2y,
      labels    = rq2y_labels_short,
      title     = rq1y_twin_labels_clean_extrashort[i],
      caption   = NULL,
      p_col     = "pval_adj",
      method    = "none"
                  )
})

patchwork::wrap_plots(
  plots,
  ncol = 3) +
  plot_annotation(
    title = "Change in Correlation",
    subtitle = expression(Cor[attritioned] ~ - ~ Cor[original]),
    theme = theme(
      plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5, size = 16, face = "bold")
    )
  )

save_plot(
  "2_rq2_2_correlations", width = 18, height = 18)
  
  

```

[📊 View Full Resolution Correlation Plots (PNG)](plots/pngs/2_rq2_2_correlations.png)


## Correlations results table

Below is a results table for the change in correlations. 

```{r}

cor_table = bind_rows(correlation_comparisons, .id = "timepoint") %>%
  mutate(
    dataset = rq1y_twin_labels_clean_extrashort[match(timepoint, rq1y_twin1)],
    x_var_label = rq2y_labels_short[match(x_var, rq2y)],
    y_var_label = rq2y_labels_short[match(y_var, rq2y)],
    pair = paste(x_var_label, "×", y_var_label)
  ) %>%
  select(dataset, pair, y, ymin, ymax, pval, pval_adj) %>%
  filter(!is.na(dataset)) %>%
  gt(groupname_col = "dataset") %>%
  # tab_options(
  #   table.width = px(800)
  # ) %>%

  # Rename columns
  cols_label(
    pair = "Variable Pair",
    y = "Est",
    ymin = "LB",
    ymax = "UB",
    pval = md("p<sub>unadj</sub>"),
    pval_adj = md("p<sub>adj</sub>")
  ) %>%

  # Create column spanner
  tab_spanner(
    label = md("Correlation difference = Cor<sub>attritioned</sub> - Cor<sub>original</sub>"),
    columns = c(y, ymin, ymax, pval, pval_adj)
  ) %>%

  # Format numbers
  fmt(
    columns = c(y, ymin, ymax),
    fns = function(x) {gbtoolbox::apa_num(x, n_decimal_places = 3)}
  ) %>%
  fmt(
    columns = c(pval, pval_adj),
    fns = function(x) {gbtoolbox::apa_num(x, n_decimal_places = 3)}
  ) %>%

  # Styling - uniform font size
  tab_style(
    style = cell_text(size = px(10)),
    locations = cells_column_spanners()
  ) %>%
  tab_style(
    style = cell_text(size = px(10)),
    locations = cells_body()
  ) %>%
  tab_style(
    style = cell_text(size = px(10), align = "center"),
    locations = cells_row_groups()
  ) %>%
  tab_style(
    style = cell_text(size = px(10)),
    locations = cells_column_labels()
  ) %>%
  tab_style(
    style = cell_text(size = px(10)),
    locations = cells_footnotes()
  ) %>%

  # Highlight significant results - positive effects (light green)
  tab_style(
    style = cell_fill(color = "#d5e8d4"),
    locations = cells_body(
      columns = c(pair, y, ymin, ymax, pval, pval_adj),
      rows = pval_adj < 0.05 & y > 0
    )
  ) %>%

  # Highlight significant results - negative effects (light red)
  tab_style(
    style = cell_fill(color = "#f8cecc"),
    locations = cells_body(
      columns = c(pair, y, ymin, ymax, pval, pval_adj),
      rows = pval_adj < 0.05 & y < 0
    )
  ) %>%

  # Add footnotes
  tab_footnote(
    footnote = md("<em>Note.</em> Est = Estimate, LB = Lower Bound 95% Confidence Interval, UB = Upper Bound 95% Confidence Interval. Significant (p<sub>Bonferroni-Holm</sub>) effects are highlighted in green (increases) or red (decreases)."),
    placement = "right"
  ) %>%
  tab_footnote(
    footnote = "P values are Bonferroni-Holm adjusted within each attrition timepoint",
    locations = cells_column_labels(columns = pval_adj),
    placement = "right"
  ) %>%
  opt_footnote_marks(marks = c("*", "†", "‡"))

cor_table

```


## Plot Original Correlations between variables

```{r}

df %>%
  filter(!(randomfamid %in% exclude_fams_onesib)) %>% 
  select(rq2y) %>%
  data.frame() %>%
  `colnames<-`(rq2y_labels_short) %>%
   gbtoolbox::plot_correlations() +
  labs(title = "Original (un-attritioned) correlations")

save_plot(
  "2_rq2_2_correlations_original", width = 8, height = 8)

```

[📊 View Plot](plots/pngs/2_rq2_2_correlations_original.png)

# ACE Comparison

## Count of significant ACE differences (unadjusted)

```{r}
bootstrap_summary_df_ace %>%
  filter(group == "Difference") %>% 
  summarise(
    total_tests = dplyr::n(),
    sig_unadj = sum(pval < 0.05),
    pct_sig = round(sig_unadj / total_tests * 100, 1)
  )
```

## ACE differences tables (p(unadjusted) < .05)

```{r}

bootstrap_summary_df_ace %>%
  filter(group == "Difference") %>%
  group_by(dataset_label) %>%
  mutate(
    pval_adj = stats::p.adjust(pval, method = "holm"),
    n_adj    = dplyr::n(),
    name     = rq2y_labels_short[match(name, rq2y)]
  ) %>%
  ungroup() %>%
  arrange(name, par, dataset_label) %>%
  filter(pval < .05) %>%
  select(-starts_with("."), -dataset, -group, -n, -pd, -n_adj) %>%
  mutate(
    par = toupper(par),
    sex = str_to_title(sex),
    row_num = row_number()
  ) %>%
  relocate(row_num, dataset_label, name, par, sex, y, ymin, ymax, pval, pval_adj) %>%
  gt() %>%
  cols_label(
    row_num = "",
    dataset_label = "Timepoint",
    name = "Variable",
    par = "Par",
    sex = "Sex",
    y = "Est",
    ymin = "LB",
    ymax = "UB",
    pval = md("p<sub>unadj</sub>"),
    pval_adj = md("p<sub>adj</sub>")
  ) %>%
  tab_spanner(
    label = "ACE Difference (Attritioned - Original)",
    columns = c(y, ymin, ymax, pval, pval_adj)
  ) %>%
  fmt(
    columns = c(y, ymin, ymax, pval, pval_adj),
    fns = function(x) {gbtoolbox::apa_num(x, n_decimal_places = 3)}
  ) %>%
  tab_style(
    style = cell_text(size = px(10)),
    locations = cells_column_spanners()
  ) %>%
  tab_style(
    style = cell_text(size = px(10)),
    locations = cells_body()
  ) %>%
  tab_style(
    style = cell_text(size = px(10)),
    locations = cells_column_labels()
  ) %>%
  tab_style(
    style = cell_text(size = px(10)),
    locations = cells_footnotes()
  ) %>%
  tab_footnote(
    footnote = md("<em>Note.</em> Est = Estimate, LB = Lower Bound 95% CI, UB = Upper Bound 95% CI, Par = ACE Parameter. Only effects with p<sub>unadjusted</sub> < .05 are shown."),
    placement = "right"
  ) %>%
  tab_footnote(
    footnote = "P values are Bonferroni-Holm adjusted within each sex, parameter, and timepoint",
    locations = cells_column_labels(columns = pval_adj),
    placement = "right"
  ) %>%
  opt_footnote_marks(marks = c("*", "†"))

```

## ACE Estimates & Differences Table (complete)

```{r}

bootstrap_summary_df_ace %>%
  filter(par %in% c("a","c","e")) %>%
  group_by(dataset_label) %>%
  mutate(
    pval = stats::p.adjust(pval, method = "holm"),
  ) %>%
  ungroup() %>%
  select(-starts_with("."), -n, -pd) %>%
  mutate(name_clean = rq2y_labels_short[match(name, rq2y)]) %>%
  pivot_wider(
    id_cols = c(par, name_clean, sex, dataset_label),
    names_from = group,
    values_from = c(y, ymin, ymax, pval),
    names_sep = "_"
  ) %>%
  select(-ymin_Original, -ymax_Original, -ymin_Attritioned, -ymax_Attritioned, -pval_Original, -pval_Attritioned) %>%
  select(dataset_label, name_clean, par, y_Original, y_Attritioned, y_Difference, ymin_Difference, ymax_Difference, pval_Difference, sex) %>%
  mutate(
    name_clean    = factor(name_clean,    levels = rq2y_labels_short),
    dataset_label = factor(dataset_label, levels = rq1y_twin_labels_clean_extrashort)
    ) %>% 
  arrange(sex, dataset_label, name_clean, par) %>%
  mutate(par = toupper(par)) %>%
  gt(groupname_col = "sex") %>%
  fmt_number(decimals = 3) %>%
  tab_row_group(
    label = "Male",
    rows = sex == "male"
  ) %>%
  tab_row_group(
    label = "Female",
    rows = sex == "female"
  ) %>%
  cols_hide(sex) %>%
  cols_label(
    dataset_label = "Timepoint",
    name_clean = "Variable",
    par = "",
    y_Original = "Original",
    y_Attritioned = "Attritioned",
    y_Difference = "Diff",
    ymin_Difference = "Lower",
    ymax_Difference = "Upper",
    pval_Difference = "p"
  ) %>%
  tab_spanner(
    label = "95% CI",
    columns = c(ymin_Difference, ymax_Difference)
  ) %>%
  tab_spanner(
    label = "Estimates",
    columns = c(y_Attritioned, y_Original)
  ) %>%
  tab_style(
    style = cell_fill(color = "lightgreen"),
    locations = cells_body(
      columns = everything(),
      rows = pval_Difference < 0.05
    )
  ) %>%
  tab_footnote(
    footnote = "P values are Bonferroni-Holm adjusted within each sex, parameter (A, C or E), and timepoint group",
    locations = cells_column_labels(columns = pval_Difference)
  )

```

## ACE Stacked Bar Plot

```{r}

bootstrap_summary_df_ace %>%
  # filter(dataset_label == "Y4") %>% 
  filter(par %in% c("a","c","e")) %>%
  filter(group != "Difference") %>%
  mutate(
    dataset_label = factor(dataset_label, levels = rq1y_twin_labels_clean_extrashort),
    name = factor(rq2y_labels_short[match(name, rq2y)], levels = rq2y_labels_short),
    group = ifelse(group == "Attritioned", "A", "O"),
    group = factor(group, levels = c("A", "O")),
    sex = str_to_title(sex),
    par = toupper(par)
  ) %>%
  ggplot(aes(x = group, y = y, fill = par)) +
  geom_col(position = "stack", alpha = 1) +
  facet_grid(dataset_label + sex ~ name, switch = "x") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = .5, size = 8),
    strip.text = element_text(size = 8),
    strip.text.x = element_text(angle = 90, hjust = 1, vjust = .5, size = 8),
    strip.text.y = element_text(angle = 0, size = 10),
    strip.placement = "outside",
    panel.grid = element_blank(),
    plot.tag = element_text(hjust = 0, vjust = 0, size = 24, face = "bold"),
    plot.tag.position = "topleft"
  ) +
  labs(
    title = "Attritioned (A) vs Original (O) ACE Estimates",
    y     = "Proportion",
    x     = NULL,
    fill  = "Component"
  ) +
  scale_fill_manual(values = c("A" = "#d73027", "C" = "#fee08b", "E" = "#4575b4")) +
  coord_cartesian(ylim = c(0, 1))

save_plot("2_rq2_ace_comparison", width = 7, height = 15, trim = TRUE)

# Create a looped version where we have plots for each timepoint

ace_plots = lapply(rq1y_twin_labels_clean_extrashort, function(timepoint) {

  bootstrap_summary_df_ace %>%
  filter(dataset_label == timepoint) %>%
  filter(par %in% c("a","c","e")) %>%
  filter(group != "Difference") %>%
  mutate(
    name = factor(rq2y_labels_short[match(name, rq2y)], levels = rq2y_labels_short),
    group = ifelse(group == "Attritioned", "A", "O"),
    group = factor(group, levels = c("A", "O")),
    sex = str_to_title(sex),
    par = toupper(par)
  ) %>%
  ggplot(aes(x = group, y = y, fill = par)) +
  geom_col(position = "stack", alpha = 1) +
  facet_grid(sex ~ name, switch = if(timepoint == "Y4") "both" else "x") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = .5, size = 8),
    axis.text.y = if(timepoint != "Y4") element_blank(),
    strip.text = element_text(size = 8),
    strip.text.x = element_text(angle = 90, hjust = 1, vjust = .5, size = 8),
    strip.text.y = element_blank(),
    strip.text.y.left = if(timepoint == "Y4") element_text(angle = 0, size = 10) else element_blank(),
    strip.placement = "outside",
    panel.grid = element_blank()
  ) +
  labs(
    title = timepoint,
    y     = if(timepoint == "Y4") "Proportion" else NULL,
    x     = NULL,
    fill  = "Component"
  ) +
  scale_fill_manual(values = c("A" = "#d73027", "C" = "#fee08b", "E" = "#4575b4")) +
  coord_cartesian(ylim = c(0, 1))
})

names(ace_plots) = rq1y_twin_labels_clean_extrashort

# Arrange all plots vertically with shared legend
patchwork::wrap_plots(ace_plots, nrow = 1) +
  patchwork::plot_layout(guides = "collect") +
  patchwork::plot_annotation(
    title = "Attritioned (A) vs Original (O) ACE Estimates by Timepoint"
  )

save_plot("2_rq2_ace_comparison_by_timepoint", width = 21, height = 9, trim = FALSE)

```

[📊 View plot 1](plots/pngs/2_rq2_ace_comparison.png)
[📊 View plot 2](plots/pngs/2_rq2_ace_comparison_by_timepoint.png)


# Supplementary Analyses 

## How predictive are our variables of later attrition?

### Correlations

```{r}

sapply(rq2y, function(x)
  sapply(rq1y_twin1, function(y)
         cor(df[[x]],df[[y]], use = "pairwise.complete.obs")
  )) %>%
  `colnames<-`(rq2y_labels_short) %>%
  `rownames<-`(rq1y_twin_labels_clean_extrashort) %>%
  knitr::kable(digits = 3)

```


### Regression Modelling 

```{r}



models = list()

for (i in seq_along(rq1y_twin1)){
  cat(i, "/", length(rq1y_twin1),"\n")
  formula <- as.formula(paste(rq1y_twin1[i], "~", paste(rq2y, collapse = "+")))
  models[[i]] = glm(formula, data = df, family = binomial, na.action = "na.omit")
}


# glm_model_comparison(models[[1]])
# 
# lapply(models, function(x) performance::r2(x))
# lapply(models, function(x) calc_auc(x))
# lapply(models, function(x) summary(x))



# calc_auc(models[[1]])

```


#### Plot of AUC and R2 for each timepoint 

```{r}

# Extract AUC and pseudo R2 for each model
model_metrics = data.frame(
  outcome = sapply(models, function(x) as.character(x$formula)[2]),
  AUC = sapply(models, calc_auc),
  R2 = sapply(models, function(x) performance::r2(x)$R2)
) %>%
  mutate(outcome = rq1y_twin_labels_clean[match(.$outcome, rq1y_twin1)])

# First, order the outcomes chronologically by extracting year numbers
model_metrics_ordered = model_metrics %>%
  mutate(
    year_num = as.numeric(str_extract(outcome, "\\d+")),
    outcome = fct_reorder(outcome, year_num)
  )

plot_data = model_metrics_ordered %>%
  pivot_longer(cols = c(AUC, R2)) %>%
  mutate(
    # Format labels
    label_text = ifelse(name == "R2", 
                       paste0(sprintf("%.1f", value * 100), "%"),
                       sprintf("%.3f", value)),
    label_text = ifelse(name == "AUC", 
                       gbtoolbox::apa_num(value, n_decimal_places = 3),
                       label_text),
    # Adjust bar start position for AUC (start from 0.5 instead of 0)
    value_start = ifelse(name == "AUC", 0.5, 0),
    value_width = value - value_start
  )

plot_data %>%
  ggplot(aes(y = outcome)) +
  geom_rect(aes(xmin = value_start, xmax = value, ymin = as.numeric(outcome) - 0.4, ymax = as.numeric(outcome) + 0.4), 
            fill = "steelblue", alpha = 0.7) + 
  geom_text(aes(x = value, label = label_text), hjust = -0.1, size = 3) +
  facet_wrap(~name, scales = "free", 
             labeller = labeller(name = c("AUC" = "AUC", "R2" = "Tjur's R²"))) +
  scale_x_continuous(
    labels = function(x) {
      # Check if we're in the R2 facet by looking at the range
      if (max(x, na.rm = TRUE) < 0.2) {  # R2 values are typically small
        paste0(round(x * 100, 1), "%")
      } else {
        sprintf("%.2f", x)
      }
    },
    expand = expansion(mult = c(0, 0.15)),
    limits = function(x) {
      # Set minimum limit to 0.5 for AUC facet (values > 0.2)
      if (max(x, na.rm = TRUE) > 0.2) {
        c(0.5, max(x, na.rm = TRUE) * 1.05)
      } else {
        c(0, max(x, na.rm = TRUE) * 1.05)
      }
    }
  ) +
  labs(
    title = "Prediction Accuracy Across Study Timepoints",
    x = "Value",
    y = "Study Timepoint"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(face = "bold"),
    plot.title = element_text(hjust = 0.5)
  )

save_plot("2_prediction_accuracy", width = 9, height = 6)

```

### Full model results 

```{r}

# Extract all coefficients and combine
models_results = do.call(rbind, lapply(1:length(models), function(i) {
  coefs = tidy(models[[i]])
  coefs$outcome = as.character(models[[i]]$formula)[2]
  return(coefs)
}))

models_results %>% 
  select(outcome, everything()) %>%
  mutate(
    outcome = var_to_label(outcome) %>% str_remove("data.*"),
    term = df_labels[match(.$term,df_colnames)],
    term   = ifelse(is.na(term), "Intercept", term),
    p_star = as.character(symnum(p.value, cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1), symbols = c("***", "**", "*", ".", " ")))
  ) %>% 
  knitr::kable(., digits = 3)


```


### Variable Importance Calculations

Note: Variable importance is calculated by measuring the change in AUC when each variable is removed from the model. 

```{r}
glm_model_comparison_results = lapply(models, glm_model_comparison)

glm_model_comparison_results_df = do.call("rbind.data.frame", glm_model_comparison_results)

# glm_model_comparison_results_df$LRT_p_star = glm_model_comparison_results_df$LRT_p %>%
#   symnum(., cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1), symbols = c("***", "**", "*", ".", " "))

# Adjust p-values and add star

glm_model_comparison_results_df = glm_model_comparison_results_df %>% 
  filter(Variables_Dropped!="None") %>%
  group_by(outcome) %>%
  mutate(LRT_p = stats::p.adjust(LRT_p, method = "holm")) %>% 
  ungroup() %>%
  mutate(
    LRT_p_star = as.character(symnum(LRT_p, cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1), symbols = c("***", "**", "*", ".", " ")))
  )

variable_importance_order = glm_model_comparison_results_df %>%
  group_by(Variables_full) %>%
  summarise(mean_score = mean(LRT_p)) %>%
  arrange(mean_score) %>%
  pull(Variables_full)
glm_model_comparison_results_df = glm_model_comparison_results_df %>%
  mutate(
    Variables_full = factor(Variables_full, levels = variable_importance_order)
  )

# Plotting code
glm_model_comparison_results_df %>%
  mutate(
    Variables_full = rq2y_labels_short[match(.$Variables_full, rq2y_labels)],
    outcome_labeled = rq1y_twin_labels_clean_extrashort[match(.$outcome, rq1y_twin1)],
    criterion       = Delta_AUC,
    sig_level = case_when(
      LRT_p < 0.001 ~ "p < 0.001",
      LRT_p < 0.01 ~ "p < 0.01", 
      LRT_p < 0.05 ~ "p < 0.05",
      LRT_p < 0.1 ~ "p < 0.1",
      TRUE ~ "Not significant"
    ),
    sig_level = factor(sig_level, levels = c("p < 0.001", "p < 0.01", "p < 0.05", "p < 0.1", "Not significant"))
  ) %>%
  ggplot(aes(y = Variables_full, x = criterion)) + 
  geom_col(aes(fill = sig_level), alpha = 0.8) + 
  geom_text(aes(label = gsub("0\\.", ".", sprintf("%.3f", criterion))),  # Remove ALL 0. patterns
            size = 2.5, fontface = "bold", color = "black") +
  facet_wrap(~outcome_labeled, scales = "fixed", ncol = 6) +
  scale_x_continuous(labels = function(x) gsub("0\\.", ".", sprintf("%.3f", x))) +  # Remove ALL 0. patterns
  scale_fill_manual(
    values = c("p < 0.001" = "#d73027", "p < 0.01" = "#fc8d59", 
               "p < 0.05" = "#fee08b", "p < 0.1" = "#e0f3f8", 
               "Not significant" = "#d9d9d9"),
    name = "Significance\n(Holm-adjusted)"
  ) +
  labs(
    x = "Change in AUC (when variable removed)",
    y = NULL,
    title = "Variable Importance by AUC Change",
    subtitle = "Colors show Holm-adjusted significance levels"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 7),
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
    strip.text = element_text(size = 9, face = "bold"),
    legend.position = "bottom"
  )


save_plot("2_rq2_variable_importance", width = 14, height = 7)
```

## Comparison of maternal education scores

```{r}

df %>% 
  # group_by(zcdata1) %>%
  summarise(
    mean = mean(amohqualn, na.rm = TRUE),
    var  = stats::var(amohqualn1, na.rm = TRUE),
    sd  = stats::sd(amohqualn1, na.rm = TRUE)
    )


df %>% 
  group_by(zcdata1) %>%
  summarise(
    mean = mean(amohqualn, na.rm = TRUE),
    var  = stats::var(amohqualn1, na.rm = TRUE),
    sd   = stats::sd(amohqualn1, na.rm = TRUE)
    
    )

```

