---
title: "RQ6: original vs attritioned"
format: 
  html:
    embed-resources: true
    fig-width: 8
    fig-height: 8
    code-link: true
    code-fold: true
    code-tools: true
    df-print: paged
    toc: true
    grid:
      body-width: 900px
editor: source
editor_options: 
  chunk_output_type: console
---

## Description

In this (non-preregistered) analysis, we look at four variables collected at years 12 & 16 which have low levels of missingness, and test the effect of selecting only those who took part at year 21 and 26 time points ("attritioning").

In scripts seven and eight, we compare the effect of attrition when using IP weighting and imputation. 

Suggested by Yasmin.

## Load Data

```{r}
#| output: false

source("0_load_data.R")

df = df %>%
  filter(!(randomfamid %in% exclude_fams_rq1x)) %>%
  filter(!(randomfamid %in% exclude_fams_onesib)) %>%
  filter(!(randomfamid %in% exclude_fams_rq6y)) 

## range of participation outcomes to use for attiritioning 

range_participation_outcomes = 6:8

variable_comparisons = readRDS(file.path("results","6_variable_comparisons.Rds"))

```


### Focal Variables and timepoints 

Attrition time points

```{r}

cbind(
  rq6y,
  rq6y_labels
) %>%
  knitr::kable(caption = "Outcome Variables")

cbind(
  rq1y_twin1[range_participation_outcomes],
  rq1y_twin_labels_clean[range_participation_outcomes]
) %>%
  knitr::kable(caption = "Attrittion time points")

```


## Data Processing

### Transform Raw Results to Dataframe

```{r}
# Load raw bootstrap results (already loaded above as variable_comparisons)
# variable_comparisons = readRDS(file.path("results", "6_variable_comparisons.Rds"))

# Extract bootstrap_summary from each element
variable_comparisons_df = lapply(variable_comparisons, function(x) x$bootstrap_summary)

# Transform each dataset's results
for(i in seq_along(variable_comparisons_df)){
  for (j in 1:3){ # Keep only md, smd, var, cor_resid (h removed from RQ6)
    variable_comparisons_df[[i]] = variable_comparisons_df[[i]][1:3]
    variable_comparisons_df[[i]][[j]] = do.call(rbind, variable_comparisons_df[[i]][[j]])
    variable_comparisons_df[[i]][[j]]$dataset  = rq1y_twin_labels_clean_extrashort[range_participation_outcomes[i]]
    variable_comparisons_df[[i]][[j]]$stat = c("md","smd","var")[j]
    variable_comparisons_df[[i]][[j]]$variable = rq6y_labels
    rownames(variable_comparisons_df[[i]][[j]]) = NULL
  }
  variable_comparisons_df[[i]] = do.call(rbind.data.frame, variable_comparisons_df[[i]])
}

# Combine all datasets into single dataframe
variable_comparisons_df0 = do.call(rbind.data.frame, variable_comparisons_df)
rownames(variable_comparisons_df0) = NULL

# Save for future use
# saveRDS(variable_comparisons_df0, file.path("results", "6_variable_comparisons_df.Rds"))

```

### Create Table

```{r}
# Get original variances for rescaling
# original_variances = sapply(rq6y, function(x) stats::var(df[[x]], na.rm = TRUE))

# variable_comparisons_df0 created above in "Transform Raw Results to Dataframe" section
# Optionally can load from disk:
# variable_comparisons_df0 = readRDS(file.path("results", "6_variable_comparisons_df.Rds"))

# Process the dataframe
variable_comparisons_df = variable_comparisons_df0 %>%
  mutate(
    # dataset = rq1y_twin_labels_clean[match(.$dataset, rq1y_twin1)],
    variable = rq6y_labels[match(.$variable, rq6y_labels)]
  ) %>%
  # Rescale variance values
  pivot_longer(cols = starts_with("y")) %>%
  # mutate(
  #   var_constant = original_variances[match(.$variable, rq6y_labels)],
  #   value = case_when(
  #     stat == "var" ~ value / var_constant,
  #     TRUE ~ value
  #   )
  # ) %>%
  # select(-var_constant) %>%
  pivot_wider(
    values_from = value,
    names_from  = name
  ) %>%
  group_by(dataset, stat) %>%
  mutate(
    pval_adj = stats::p.adjust(pval, method = "holm")
  )

```

## Means, SMDs, Variances

### GT Results Table

```{r}

variable_comparisons_df %>%
  filter(stat %in% c("md", "smd", "var")) %>%
  select(-starts_with("."),-pd) %>%
  pivot_wider(
    names_from = "stat",
    values_from = c("y","ymin","ymax","pval","pval_adj"),
    id_cols = c("variable", "dataset")
  ) %>%
  dplyr::rename(Outcome = "variable") %>%
  gt(groupname_col = "dataset") %>%
  tab_options(
    table.width = px(800)
  ) %>%

  # Create column spanners (nested headers)
  tab_spanner(
    label = "Mean Difference",
    columns = c(y_md, ymin_md, ymax_md, pval_md, pval_adj_md)
  ) %>%
  tab_spanner(
    label = "Standardized Mean Difference",
    columns = c(y_smd, ymin_smd, ymax_smd, pval_smd, pval_adj_smd)
  ) %>%
  tab_spanner(
    label = "Variance % Change",
    columns = c(y_var, ymin_var, ymax_var, pval_var, pval_adj_var)
  ) %>%

  # Rename columns under each spanner
  cols_label(
    y_md = "Est", ymin_md = "LB", ymax_md = "UB", pval_md = md("p<sub>unadj</sub>"), pval_adj_md = md("p<sub>adj</sub>"),
    y_smd = "Est", ymin_smd = "LB", ymax_smd = "UB", pval_smd = md("p<sub>unadj</sub>"), pval_adj_smd = md("p<sub>adj</sub>"),
    y_var = "Est", ymin_var = "LB", ymax_var = "UB", pval_var = md("p<sub>unadj</sub>"), pval_adj_var = md("p<sub>adj</sub>")
  ) %>%

  # Format numbers
  fmt(
    columns = !contains("var") & !contains("Outcome"),
    fns = function(x) {gbtoolbox::apa_num(x, n_decimal_places = 3)}
  ) %>%
  fmt(
    columns = c("pval_var", "pval_adj_var"),
    fns = function(x) {gbtoolbox::apa_num(x, n_decimal_places = 3)}
  ) %>%
  fmt_percent(
    columns = c("y_var", "ymin_var","ymax_var"),
    decimals = 2,
    drop_trailing_zeros = FALSE,
    drop_trailing_dec_mark = FALSE
  ) %>%
  # Styling - uniform font size
  tab_style(
    style = cell_text(size = px(10)),
    locations = cells_column_spanners()
  ) %>%
  tab_style(
    style = cell_text(size = px(10)),
    locations = cells_body()
  ) %>%
  tab_style(
    style = cell_text(size = px(10), align = "center"),
    locations = cells_row_groups()
  ) %>%
  tab_style(
    style = cell_text(size = px(10)),
    locations = cells_column_labels()
  ) %>%
  tab_style(
    style = cell_text(size = px(10)),
    locations = cells_footnotes()
  ) %>%

  # Highlight significant results - positive effects (light green)
  tab_style(
    style = cell_fill(color = "#d5e8d4"),
    locations = cells_body(
      columns = c(y_md, ymin_md, ymax_md, pval_md, pval_adj_md),
      rows = pval_adj_md < 0.05 & y_md > 0
    )
  ) %>%
  tab_style(
    style = cell_fill(color = "#d5e8d4"),
    locations = cells_body(
      columns = c(y_smd, ymin_smd, ymax_smd, pval_smd, pval_adj_smd),
      rows = pval_adj_smd < 0.05 & y_smd > 0
    )
  ) %>%
  tab_style(
    style = cell_fill(color = "#d5e8d4"),
    locations = cells_body(
      columns = c(y_var, ymin_var, ymax_var, pval_var, pval_adj_var),
      rows = pval_adj_var < 0.05 & y_var > 0
    )
  ) %>%

  # Highlight significant results - negative effects (light red)
  tab_style(
    style = cell_fill(color = "#f8cecc"),
    locations = cells_body(
      columns = c(y_md, ymin_md, ymax_md, pval_md, pval_adj_md),
      rows = pval_adj_md < 0.05 & y_md < 0
    )
  ) %>%
  tab_style(
    style = cell_fill(color = "#f8cecc"),
    locations = cells_body(
      columns = c(y_smd, ymin_smd, ymax_smd, pval_smd, pval_adj_smd),
      rows = pval_adj_smd < 0.05 & y_smd < 0
    )
  ) %>%
  tab_style(
    style = cell_fill(color = "#f8cecc"),
    locations = cells_body(
      columns = c(y_var, ymin_var, ymax_var, pval_var, pval_adj_var),
      rows = pval_adj_var < 0.05 & y_var < 0
    )
  ) %>%

  # Add footnotes
  tab_footnote(
    footnote = md("<em>Note.</em> Est = Estimate, LB = Lower Bound 95% Confidence Interval, UB = Upper Bound 95% Confidence Interval. Significant (p<sub>Bonferroni-Holm</sub>) effects are highlighted in green (increases) or red (decreases)."),
    placement = "right"
  ) %>%
  tab_footnote(
    footnote = "P values are Bonferroni-Holm adjusted within each participation group",
    locations = cells_column_labels(columns = contains("pval_adj")),
    placement = "right"
  ) %>%
  tab_footnote(
    footnote = md("Standardized mean difference = (Mean<sub>original</sub> - Mean<sub>attritioned</sub>) / SD<sub>original</sub>"),
    locations = cells_column_spanners(spanners = "Standardized Mean Difference"),
    placement = "right"
  ) %>%
  tab_footnote(
    footnote = md("Percentage change = (Var<sub>attritioned</sub> - Var<sub>original</sub>) / Var<sub>original</sub>"),
    locations = cells_column_spanners(spanners = "Variance % Change"),
    placement = "right"
  ) %>%
  opt_footnote_marks(marks = c("*", "†","‡"))

```

### Create Circular Heatmaps

#### Standardised Mean Differences

```{r}
# Create circular heatmap for SMD
heatmap_data_smd = variable_comparisons_df %>%
  filter(stat == "smd") %>%
  mutate(
    outcome_labeled = rq1y_twin_labels_clean_extrashort[match(dataset, rq1y_twin_labels_clean_extrashort)],
    outcome_labeled = factor(outcome_labeled, levels = rq1y_twin_labels_clean_extrashort[range_participation_outcomes]),
    Variables_wrapped = str_wrap(variable, width = 8),
    # Set fill value to NA if not significant, otherwise keep original sign
    fill_value = case_when(
      pval_adj < 0.05 ~ y,
      TRUE ~ NA
    )
  ) %>%
  # Use fixed order from rq6y_labels
  mutate(
    variable = factor(variable, levels = rq6y_labels),
    Variables_wrapped = factor(Variables_wrapped, levels = str_wrap(rq6y_labels, width = 8))
  )

ggplot(heatmap_data_smd, aes(x = Variables_wrapped, y = outcome_labeled)) +
  geom_tile(aes(fill = fill_value),
            color = "black",
            size = 0.5) +
  geom_text(
    aes(label = ifelse(pval_adj < 0.05,
                       gsub("^(-?)0\\.", "\\1.", sprintf("%.3f", y)), ""),
        size = 6.5
    ),
    color = "black"
  ) +
  scale_size_identity() +
  coord_polar(theta = "x", start = 0, direction = 1, clip = "off") +
  scale_fill_gradient2(
    low = "#2166ac",
    mid = "white",
    high = "#d73027",
    midpoint = 0,
    limits = c(-0.3, 0.3),
    na.value = "white",
    guide = "none"
  ) +
  labs(
    subtitle = expression(frac(bar(X)[attritioned] - bar(X)[original], SD[original])),
    tag = "A1"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 0, size = 13, color = "black"),
    axis.text.y = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    plot.title = element_blank(),
    plot.subtitle = element_text(hjust = 0.5, size = 17, margin = margin(b = -45)),
    plot.tag = element_text(hjust = 0, vjust = 0, size = 30, face = "bold"),
    plot.tag.position = "topleft"
  ) +
  {
    y_positions = seq_along(levels(heatmap_data_smd$outcome_labeled))
    labels = levels(heatmap_data_smd$outcome_labeled)

    lapply(seq_along(labels), function(i) {
      annotate("text", x = 0.5, y = y_positions[i],
               label = paste0(labels[i], " "), size = 5, hjust = 1, angle = 7, vjust = -.3, fontface = "bold")
    })
  }

save_plot("6_rq6_circular_heatmap_smd", width = 8, height = 8, trim = TRUE)

```

[📊 View Full Resolution SMD Heatmap (PNG)](plots/pngs/6_rq6_circular_heatmap_smd.png)

#### Variance Differences

```{r}
# Create circular heatmap for variance differences
heatmap_data_var = variable_comparisons_df %>%
  filter(stat == "var") %>%
  mutate(
    outcome_labeled = rq1y_twin_labels_clean_extrashort[match(dataset, rq1y_twin_labels_clean_extrashort)],
    outcome_labeled = factor(outcome_labeled, levels = rq1y_twin_labels_clean_extrashort[range_participation_outcomes]),
    Variables_wrapped = str_wrap(variable, width = 8),
    # Set fill value to NA if not significant, otherwise keep original sign
    fill_value = case_when(
      pval_adj < 0.05 ~ y,
      TRUE ~ NA
    )
  ) %>%
  # Use fixed order from rq6y_labels
  mutate(
    variable = factor(variable, levels = rq6y_labels),
    Variables_wrapped = factor(Variables_wrapped, levels = str_wrap(rq6y_labels, width = 8))
  )

ggplot(heatmap_data_var, aes(x = Variables_wrapped, y = outcome_labeled)) +
  geom_tile(aes(fill = fill_value),
            color = "black",
            size = 0.5) +
  geom_text(
    aes(label = ifelse(pval_adj < 0.05,
                       paste0(sprintf("%.1f", y * 100), "%"), ""),
        size = 6.5
    ),
    color = "black"
  ) +
  scale_size_identity() +
  coord_polar(theta = "x", start = 0, direction = 1, clip = "off") +
  scale_fill_gradient2(
    low = "#2166ac",
    mid = "white",
    high = "#d73027",
    midpoint = 0,
    limits = c(-.135, .135),
    na.value = "white",
    guide = "none"
  ) +
  labs(
    subtitle = expression(frac(Var[attritioned] - Var[original], Var[original]) ~ "\u00D7" ~ 100),
    tag = "B1"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 0, size = 13, color = "black"),
    axis.text.y = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    plot.title = element_blank(),
    plot.subtitle = element_text(hjust = 0.5, size = 17, margin = margin(b = -45)),
    plot.tag = element_text(hjust = 0, vjust = 0, size = 30, face = "bold"),
    plot.tag.position = "topleft"
  ) +
  {
    y_positions = seq_along(levels(heatmap_data_var$outcome_labeled))
    labels = levels(heatmap_data_var$outcome_labeled)

    lapply(seq_along(labels), function(i) {
      annotate("text", x = 0.5, y = y_positions[i],
               label = paste0(labels[i], " "), size = 5, hjust = 1, angle = 7, vjust = -.3, fontface = "bold")
    })
  }

save_plot("6_rq6_circular_heatmap_var", width = 8, height = 8, trim = TRUE)

```

[📊 View Full Resolution Variance Heatmap (PNG)](plots/pngs/6_rq6_circular_heatmap_var.png)

## Correlations

First, some code to establish how to convert variable labels from matrix to dataframe formats.

```{r}
test_correlation_matrix = matrix(
  nrow = length(rq6y),
  ncol = length(rq6y)
  )

for(i in seq_along(rq6y)){
  for(j in seq_along(rq6y)){
    test_correlation_matrix[i,j] = paste(rq6y[i], rq6y[j], collapse = " ")
  }
}

x = test_correlation_matrix[lower.tri(test_correlation_matrix)]

x_var = str_extract(x, "^\\S+")
y_var = str_extract(x, "\\S+$")

rm(test_correlation_matrix,x)
```

### Plot of differences in correlations

```{r}

correlation_comparisons = lapply(variable_comparisons, function(x) x$bootstrap_summary$cor_resid)

correlation_comparisons = lapply(correlation_comparisons, function(x) do.call(rbind.data.frame, x))

correlation_comparisons = lapply(correlation_comparisons, function(x) {
  x %>%
    mutate(
      pval_adj = stats::p.adjust(pval, method = "holm"),
      x_var = x_var,
      y_var = y_var
    )
})

library(patchwork)

plots = lapply(1:length(correlation_comparisons), function(i) {
  plot_lower_triangular_matrix2(
    data = correlation_comparisons[[i]],
    variables = rq6y,
    labels = rq6y_labels,
    x_var_col = "x_var",
    y_var_col = "y_var",
    value_col = "y",
    p_col = "pval_adj",
    p_threshold = 0.05,
    show_values = TRUE,
    text_size = 3,
    title = rq1y_twin_labels_clean_extrashort[range_participation_outcomes[i]],
    caption = NULL,
    method = "none"
  ) +
  labs(tag = paste0("A", i)) +
  theme(
    plot.title        = element_text(hjust=0),
    axis.text         = element_text(color = "black"),
    plot.background   = element_blank(),
    panel.background  = element_blank(),
    legend.background = element_blank(),
    plot.tag          = element_text(size = 14, face = "bold"),
    plot.tag.position = c(0, 1)
  )
})

patchwork::wrap_plots(
  plots,
  ncol = 3) +
  plot_annotation(
    title = expression(Cor[attritioned] ~ - ~ ~ Cor[original]),
    theme = theme(
      plot.title    = element_text(hjust = 0.5, size = 16, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5, size = 13)
      )
  )

save_plot(
  "6_rq6_correlations", width = 11, height = 3, trim = TRUE, trim_margin = 0
  )



```

[📊 View Full Resolution Plot](plots/pdfs/6_rq6_correlations.png)

### GT Results table

```{r}

cor_table = bind_rows(correlation_comparisons, .id = "timepoint") %>%
  mutate(
    dataset = rq1y_twin_labels_clean_extrashort[match(timepoint, rq1y_twin1)],
    x_var_label = rq6y_labels[match(x_var, rq6y)],
    y_var_label = rq6y_labels[match(y_var, rq6y)],
    pair = paste(x_var_label, "×", y_var_label)
  ) %>%
  select(dataset, pair, y, ymin, ymax, pval, pval_adj) %>%
  gt(groupname_col = "dataset") %>%
  tab_options(
    table.width = px(800)
  ) %>%

  # Rename columns
  cols_label(
    pair = "Variable Pair",
    y = "Est",
    ymin = "LB",
    ymax = "UB",
    pval = md("p<sub>unadj</sub>"),
    pval_adj = md("p<sub>adj</sub>")
  ) %>%

  # Create column spanner
  tab_spanner(
    label = "Correlation Difference",
    columns = c(y, ymin, ymax, pval, pval_adj)
  ) %>%

  # Format numbers
  fmt(
    columns = c(y, ymin, ymax),
    fns = function(x) {gbtoolbox::apa_num(x, n_decimal_places = 3)}
  ) %>%
  fmt(
    columns = c(pval, pval_adj),
    fns = function(x) {gbtoolbox::apa_num(x, n_decimal_places = 3)}
  ) %>%

  # Styling - uniform font size
  tab_style(
    style = cell_text(size = px(10)),
    locations = cells_column_spanners()
  ) %>%
  tab_style(
    style = cell_text(size = px(10)),
    locations = cells_body()
  ) %>%
  tab_style(
    style = cell_text(size = px(10), align = "center"),
    locations = cells_row_groups()
  ) %>%
  tab_style(
    style = cell_text(size = px(10)),
    locations = cells_column_labels()
  ) %>%
  tab_style(
    style = cell_text(size = px(10)),
    locations = cells_footnotes()
  ) %>%

  # Highlight significant results - positive effects (light green)
  tab_style(
    style = cell_fill(color = "#d5e8d4"),
    locations = cells_body(
      columns = c(y, ymin, ymax, pval, pval_adj),
      rows = pval_adj < 0.05 & y > 0
    )
  ) %>%

  # Highlight significant results - negative effects (light red)
  tab_style(
    style = cell_fill(color = "#f8cecc"),
    locations = cells_body(
      columns = c(y, ymin, ymax, pval, pval_adj),
      rows = pval_adj < 0.05 & y < 0
    )
  ) %>%

  # Add footnotes
  tab_footnote(
    footnote = md("<em>Note.</em> Est = Estimate, LB = Lower Bound 95% Confidence Interval, UB = Upper Bound 95% Confidence Interval. Significant (p<sub>Bonferroni-Holm</sub>) effects are highlighted in green (increases) or red (decreases)."),
    placement = "right"
  ) %>%
  tab_footnote(
    footnote = "P values are Bonferroni-Holm adjusted within each attrition timepoint",
    locations = cells_column_labels(columns = pval_adj),
    placement = "right"
  ) %>%
  tab_footnote(
    footnote = md("Correlation difference = Cor<sub>attritioned</sub> - Cor<sub>original</sub>"),
    locations = cells_column_spanners(spanners = "Correlation Difference"),
    placement = "right"
  ) %>%
  opt_footnote_marks(marks = c("*", "†", "‡"))

cor_table

```

### Percentiles of effects

```{r}
bind_rows(correlation_comparisons, .id = "timepoint") %>%
  mutate(
    x_var_label = rq6y_labels[match(x_var, rq6y)],
    y_var_label = rq6y_labels[match(y_var, rq6y)],
    pair = paste(x_var_label, "×", y_var_label)
  ) %>%
  select(pair, y, pval_adj) %>%
  summarise(
    n_pairs = dplyr::n(),
    mean    = mean(abs(y)),
    p50     = quantile(abs(y), .50),
    p90     = quantile(abs(y), .90),
    p95     = quantile(abs(y), .95),
    max     = max(abs(y))
  ) %>%
  gt() %>%
  cols_label(
    n_pairs = "Pairs (total)",
    mean    = "Mean",
    p50     = "Median",
    p90     = "90th",
    p95     = "95th",
    max     = "Max"
  ) %>%
  fmt_number(columns = c(mean, p50, p90, p95, max), decimals = 3) %>%
  tab_spanner(
    label = md("|Correlation Change| Percentiles"),
    columns = c(p50, p90, p95, max)
  ) %>%
  tab_style(
    style = cell_text(size = px(11)),
    locations = list(cells_body(), cells_column_labels(), cells_column_spanners())
  ) %>%
  tab_footnote(
    footnote = md("<em>Note.</em> Percentiles of |r<sub>attritioned</sub> − r<sub>original</sub>| across all variable pairs and timepoints."),
    placement = "right"
  )
```

### Plot of original correlations

```{r}

df %>%
  select(rq6y) %>%
  `colnames<-`(rq6y_labels) %>%
  as.data.frame() %>%
  gbtoolbox::plot_correlations(
    confidence_interval = FALSE,
    text_size_r = 5
  ) +
  labs(
    title = "Original (unattritioned) correlations",
    subtitle = "**Below Diagonal:** Correlations.<br>**Diagonal:** Univariate Sample Sizes.<br>**Above Diagonal:** Pairwise Sample Sizes") +
  theme(
    axis.text = element_text(color = "black", size = 10),
    plot.subtitle = ggtext::element_markdown()
  )

save_plot("6_rq6_original_correlations", height = 5.2, width = 5.2)
```

[View Plot](/plots/pngs/6_rq6_original_correlations.png)

## ACE Comparison

```{r}

# Extract ACE estimates and differences from all datasets
ace_estimates_list = lapply(variable_comparisons, function(x) x$ace_est_summary)
ace_differences_list = lapply(variable_comparisons, function(x) x$ace_diff_summary)

# Add dataset identifier to each
for(i in seq_along(ace_estimates_list)){
  ace_estimates_list[[i]]$dataset = rq1y_twin1[range_participation_outcomes[i]]
  ace_differences_list[[i]]$dataset = rq1y_twin1[range_participation_outcomes[i]]
}

# Combine into single dataframes
ace_estimates = do.call(rbind, ace_estimates_list)
ace_differences = do.call(rbind, ace_differences_list)

# Add readable dataset labels
ace_estimates$dataset_label = rq1y_twin_labels_clean[match(ace_estimates$dataset, rq1y_twin1)]
ace_differences$dataset_label = rq1y_twin_labels_clean[match(ace_differences$dataset, rq1y_twin1)]

```

### Table - ACE estimates

```{r}

ace_estimates %>%
  filter(par %in% c("a", "c", "e")) %>%
  select(-starts_with("."), -n, -pd, -pval) %>%
  mutate(
    name_clean = rq6y_labels[match(name, rq6y)],
    group = ifelse(group == "df1", "Attritioned", "Original")
  ) %>%
  pivot_wider(
    id_cols = c(dataset_label, name_clean, par, sex),
    names_from = group,
    values_from = c(y, ymin, ymax),
    names_sep = "_"
  ) %>%
  select(dataset_label, name_clean, par, sex,
         y_Original, ymin_Original, ymax_Original,
         y_Attritioned, ymin_Attritioned, ymax_Attritioned) %>%
  arrange(sex, dataset_label, name_clean, par) %>%
  mutate(par = toupper(par)) %>%
  gt(groupname_col = "sex") %>%
  fmt_number(decimals = 3) %>%
  tab_row_group(
    label = "Male",
    rows = sex == "male"
  ) %>%
  tab_row_group(
    label = "Female",
    rows = sex == "female"
  ) %>%
  cols_hide(sex) %>%
  cols_label(
    dataset_label = "Dataset",
    name_clean = "Variable",
    par = "",
    y_Original = "Est",
    ymin_Original = "Lower",
    ymax_Original = "Upper",
    y_Attritioned = "Est",
    ymin_Attritioned = "Lower",
    ymax_Attritioned = "Upper"
  ) %>%
  tab_spanner(
    label = "Original",
    columns = c(y_Original, ymin_Original, ymax_Original)
  ) %>%
  tab_spanner(
    label = "Attritioned",
    columns = c(y_Attritioned, ymin_Attritioned, ymax_Attritioned)
  ) %>%
  tab_options(
    table.font.size = px(9)
  )

```

### Table - Changes in ACE Estimates

```{r}

ace_differences_tbl = ace_differences  %>%
  group_by(dataset) %>%
  mutate(
    pval_adj = stats::p.adjust(pval, method = "holm")
  ) %>%
  ungroup()

ace_differences_tbl %>%
  select(-starts_with("."), -n, -pd) %>%
  mutate(name_clean = rq6y_labels[match(name, rq6y)]) %>%
  select(dataset_label, name_clean, par, sex, y, ymin, ymax, pval, pval_adj) %>%
  arrange(sex, dataset_label, name_clean, par) %>%
  mutate(par = toupper(par)) %>%
  gt(groupname_col = "sex") %>%
  fmt_number(decimals = 3) %>%
  tab_row_group(
    label = "Male",
    rows = sex == "male"
  ) %>%
  tab_row_group(
    label = "Female",
    rows = sex == "female"
  ) %>%
  cols_hide(sex) %>%
  cols_label(
    dataset_label = "Dataset",
    name_clean = "Variable",
    par = "",
    y = "Diff",
    ymin = "Lower",
    ymax = "Upper",
    pval = md("p<sub>unadj</sub>"),
    pval_adj = md("p<sub>adj</sub>")
  ) %>%
  tab_spanner(
    label = "95% CI",
    columns = c(ymin, ymax)
  ) %>%
  tab_style(
    style = cell_fill(color = "#d5e8d4"),
    locations = cells_body(
      columns = everything(),
      rows = pval_adj < 0.05 & y > 0
    )
  ) %>%
  tab_style(
    style = cell_fill(color = "#f8cecc"),
    locations = cells_body(
      columns = everything(),
      rows = pval_adj < 0.05 & y < 0
    )
  ) %>%
  tab_footnote(
    footnote = "Difference = Attritioned − Original. Adjusted P values are Bonferroni-Holm corrected within each sex, dataset, and parameter (A, C or E) group",
    locations = cells_column_labels(columns = pval_adj)
  ) %>%
  tab_options(
    table.font.size = px(9)
  )

```

### Percentiles of effects

```{r}
ace_differences_tbl %>%
  select(y) %>%
  summarise(
    n_tests = dplyr::n(),
    mean    = mean(abs(y)),
    p50     = quantile(abs(y), .50),
    p90     = quantile(abs(y), .90),
    p95     = quantile(abs(y), .95),
    max     = max(abs(y))
  ) %>%
  gt() %>%
  cols_label(
    n_tests = "Tests (total)",
    mean    = "Mean",
    p50     = "Median",
    p90     = "90th",
    p95     = "95th",
    max     = "Max"
  ) %>%
  fmt_number(columns = c(mean, p50, p90, p95, max), decimals = 3) %>%
  tab_spanner(
    label = md("|ACE Difference| Percentiles"),
    columns = c(p50, p90, p95, max)
  ) %>%
  tab_style(
    style = cell_text(size = px(11)),
    locations = list(cells_body(), cells_column_labels(), cells_column_spanners())
  ) %>%
  tab_footnote(
    footnote = md("<em>Note.</em> Percentiles of |ACE<sub>attritioned</sub> − ACE<sub>original</sub>| across all variables, timepoints, and ACE components (A, C, E)."),
    placement = "right"
  )
```

### Directional Consistency Across Variables

```{r}
# For each ACE component × variable, how many of the 6 sex × timepoint
# combinations show a positive vs negative change, and how many are significant?
ace_differences_tbl %>%
  mutate(
    par        = toupper(par),
    name_clean = rq6y_labels[match(name, rq6y)]
  ) %>%
  group_by(par, name_clean) %>%
  summarise(
    n          = dplyr::n(),
    n_pos      = sum(y > 0),
    n_neg      = sum(y < 0),
    n_sig_pos  = sum(pval < 0.05 & y > 0),
    n_sig_neg  = sum(pval < 0.05 & y < 0),
    mean_diff  = mean(y),
    .groups    = "drop"
  ) %>%
  arrange(par, name_clean) %>%
  gt(groupname_col = "par", rowname_col = "name_clean") %>%
  cols_label(
    n          = "N",
    n_pos      = md("N &uarr;"),
    n_neg      = md("N &darr;"),
    n_sig_pos  = md("N sig &uarr;"),
    n_sig_neg  = md("N sig &darr;"),
    mean_diff  = "Mean diff"
  ) %>%
  fmt_number(columns = mean_diff, decimals = 3) %>%
  tab_style(
    style = cell_text(size = px(10)),
    locations = list(
      cells_body(), cells_stub(), cells_row_groups(), cells_column_labels()
    )
  ) %>%
  tab_style(
    style = cell_fill(color = "#d5e8d4"),
    locations = cells_body(columns = n_pos, rows = n_pos == n)
  ) %>%
  tab_style(
    style = cell_fill(color = "#f8cecc"),
    locations = cells_body(columns = n_neg, rows = n_neg == n)
  ) %>%
  tab_footnote(
    footnote = md("<em>Note.</em> N = number of sex × timepoint combinations (max 6). N &uarr; / N &darr; = number with positive/negative change (Attritioned − Original). N sig = number with p<sub>unadj</sub> < 0.05. Cells highlighted when all combinations agree in direction."),
    placement = "right"
  )
```

### Forest Plot of ACE Changes

```{r}
ace_differences_tbl %>%
  mutate(
    par           = toupper(par),
    sex           = str_to_title(sex),
    name_clean    = rq6y_labels[match(name, rq6y)],
    name_clean    = factor(name_clean, levels = rev(rq6y_labels)),
    dataset_label = factor(
      rq1y_twin_labels_clean_extrashort[match(dataset, rq1y_twin1)],
      levels = rq1y_twin_labels_clean_extrashort[range_participation_outcomes]
    ),
    significant   = pval_adj < 0.05
  ) %>%
  ggplot(aes(
    x     = y,
    xmin  = ymin,
    xmax  = ymax,
    y     = name_clean,
    color = dataset_label,
    alpha = significant
  )) +
  geom_pointrange(position = position_dodge(width = 0.6), size = 0.4, linewidth = 0.7) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40", linewidth = 0.5) +
  facet_grid(sex ~ par, scales = "free_x") +
  scale_color_brewer(palette = "Dark2", name = "Attrition timepoint") +
  scale_alpha_manual(values = c("TRUE" = 1, "FALSE" = 0.35), guide = "none") +
  labs(
    title    = "Changes in ACE Estimates by Variable",
    subtitle = "Attritioned − Original (95% CI). Dashed line = no change. Faded = non-significant (p\u2090\u2090\u2090 \u2265 0.05).",
    x        = "ACE Difference",
    y        = NULL
  ) +
  theme_bw() +
  theme(
    strip.text      = element_text(face = "bold", size = 10),
    axis.text       = element_text(color = "black", size = 9),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )

save_plot("6_rq6_ace_forest", width = 7, height = 5)
```

### ACE Estimates Plot

```{r}

ace_estimates %>%
  mutate(
    name_clean = rq6y_labels[match(name, rq6y)],
    name_clean = factor(name_clean, levels = rq6y_labels),
    group = ifelse(group == "df1", "A", "O"),
    group = factor(group, levels = c("A", "O")),
    sex = str_to_title(sex),
    par = toupper(par),
    dataset_label = factor(rq1y_twin_labels_clean_extrashort[match(dataset,rq1y_twin1)], levels = rq1y_twin_labels_clean_extrashort[range_participation_outcomes])
  ) %>%
  ggplot(aes(x = group, y = y, fill = par)) +
  geom_col(position = "stack", alpha = 1) +
  facet_grid(sex + dataset_label ~ name_clean, switch = "x") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = .5, size = 8),
    strip.text = element_text(size = 8),
    strip.text.x = element_text(angle = 90, hjust = 1, vjust = .5),
    strip.text.y = element_text(angle = 0),
    strip.placement = "outside",
    panel.grid = element_blank(),
    plot.tag = element_text(hjust = 0, vjust = 0, size = 24, face = "bold"),
    plot.tag.position = "topleft"
  ) +
  labs(
    title = "Attritioned vs Original",
    x = NULL,
    y = "Proportion",
    fill = "Component",
    tag = "A"
  ) +
  scale_y_continuous(breaks = c(.25,.50,.75, 1), labels = c(".25",".50",".75","1.00")) +
  scale_fill_manual(values = c("A" = "#d73027", "C" = "#fee08b", "E" = "#4575b4")) +
  coord_cartesian(ylim = c(0, 1))

save_plot("6_rq6_ace_comparison", width = 5, height = 8)

```

[📊 View Full Resolution ACE Comparison Plot (PNG)](plots/pngs/6_rq6_ace_comparison.png)

Although we see large-ish increases in the shared environment for depression symptoms in males, this is not significant even before multiple comparisons are corrected for. 

## Effect of attritioning on sample sizes

```{r}

response_statistics_rq6 = df %>%
  select(randomtwinid, all_of(rq6y), all_of(rq1y_twin1[range_participation_outcomes])) %>%
  mutate(all1 = 1) %>%
  pivot_longer(cols = all_of(rq6y), names_to = "variable", values_to = "score") %>%
  pivot_longer(cols = all_of(c("all1", rq1y_twin1[range_participation_outcomes])), names_to = "timepoint", values_to = "participation_indicator") %>%
  filter(!is.na(score)) %>%
  filter(participation_indicator == 1) %>%
  group_by(variable, timepoint) %>%
  summarise(
    n = dplyr::n(),
    .groups = "drop"
  ) %>%
  group_by(variable) %>%
  mutate(
    max     = base::max(n),
    percent = n / max
  )

response_statistics_rq6 %>%
  mutate(
    variable  = rq6y_labels[match(variable, rq6y)],
    variable  = factor(variable, levels = rq6y_labels),
    timepoint = rq1y_twin_labels_clean_extrashort[match(timepoint, rq1y_twin1)],
    timepoint = ifelse(is.na(timepoint), "Original", timepoint),
    timepoint = factor(timepoint, levels = c("Original", rq1y_twin_labels_clean_extrashort[range_participation_outcomes])),
    original  = factor(ifelse(timepoint == "Original", "Original", "Attritioned"))
  ) %>%
  ggplot(aes(x = timepoint, y = n, group = variable)) +
  geom_line() +
  geom_point(aes(shape = original, size = original)) +
  geom_text(
    data = ~ filter(.x, timepoint == "Original"),
    aes(label = paste0("N = ", as.character(n))),
    vjust = .5, hjust = 1.2, size = 2.2, angle = 90, fontface = "bold"
  ) +
  geom_text(
    data = ~ filter(.x, timepoint != "Original"),
    aes(label = paste0(round(percent * 100, digits = 1), "%")),
    vjust = .5, hjust = -.35, size = 1.9, angle = 90
  ) +
  scale_y_continuous(limits = c(0, NA)) +
  scale_shape_manual(values = c("Original" = 18, "Attritioned" = 16)) +
  scale_size_manual(values = c("Original" = 3.5, "Attritioned" = 2)) +
  labs(
    y     = "Sample size post-attritioning",
    x     = NULL,
    title = "Adolescence Measures"
  ) +
  facet_wrap(~variable, nrow = 1) +
  theme_bw() +
  theme(
    axis.text   = element_text(color = "black"),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 7),
    legend.position = "none"
  )

save_plot("6_rq6_attrition_numbers", width = 7, height = 2.2)

```

[📊 View Full Resolution Attrition Numbers Plot (PNG)](plots/pngs/6_rq6_attrition_numbers.png)

```{r}

ace_differences_tbl %>%
  filter(sex == "male" & name == "lcmfqt1") %>%
  select(-starts_with("."), -n, -pd) %>%
  mutate(name_clean = rq6y_labels[match(name, rq6y)]) %>%
  select(dataset_label, name_clean, par, sex, y, ymin, ymax, pval, pval_adj) %>%
  arrange(sex, dataset_label, name_clean, par) %>%
  mutate(par = toupper(par)) %>%
  gt(groupname_col = "sex") %>%
  fmt_number(decimals = 3) %>%
  tab_row_group(
    label = "Male",
    rows = sex == "male"
  ) %>%
  tab_row_group(
    label = "Female",
    rows = sex == "female"
  ) %>%
  cols_hide(sex) %>%
  cols_label(
    dataset_label = "Dataset",
    name_clean = "Variable",
    par = "",
    y = "Diff",
    ymin = "Lower",
    ymax = "Upper",
    pval = md("p<sub>unadj</sub>"),
    pval_adj = md("p<sub>adj</sub>")
  ) %>%
  tab_spanner(
    label = "95% CI",
    columns = c(ymin, ymax)
  ) %>%
  tab_style(
    style = cell_fill(color = "#d5e8d4"),
    locations = cells_body(
      columns = everything(),
      rows = pval_adj < 0.05 & y > 0
    )
  ) %>%
  tab_style(
    style = cell_fill(color = "#f8cecc"),
    locations = cells_body(
      columns = everything(),
      rows = pval_adj < 0.05 & y < 0
    )
  ) %>%
  tab_footnote(
    footnote = "Difference = Attritioned − Original. Adjusted P values are Bonferroni-Holm corrected within each sex, dataset, and parameter (A, C or E) group",
    locations = cells_column_labels(columns = pval_adj)
  ) %>%
  tab_options(
    table.font.size = px(9)
  )

```

