---
title: "RQ6: Combined Analysis of Bias Across Conditions"
format: 
  html:
    embed-resources: true
    fig-width: 8
    fig-height: 8
    code-link: true
    code-fold: true
    code-tools: true
    df-print: paged
    toc: true
    grid:
      body-width: 900px
editor: source
editor_options: 
  chunk_output_type: console
---

# Load & Clean Data

```{r}
#| output: false

source("0_load_data.R")

df = df %>%
  filter(!(randomfamid %in% exclude_fams_rq1x)) %>%
  filter(!(randomfamid %in% exclude_fams_onesib)) %>%
  filter(!(randomfamid %in% exclude_fams_rq6y))

## range of participation outcomes to use for attrition

range_participation_outcomes = 6:8

# Subset vectors to only include participation outcomes in range
rq1y_twin                         = rq1y_twin[range_participation_outcomes]
rq1y_twin1                        = rq1y_twin1[range_participation_outcomes]
rq1y_twin2                        = rq1y_twin2[range_participation_outcomes]
rq1y_twin_labels                  = rq1y_twin_labels[range_participation_outcomes]
rq1y_twin_labels_clean            = rq1y_twin_labels_clean[range_participation_outcomes]
rq1y_twin_labels_clean_extrashort = rq1y_twin_labels_clean_extrashort[range_participation_outcomes]

## range of participation outcomes to use for attiritioning 

range_participation_outcomes = 6:8

res_unadj = readRDS(file = file.path("results", "6_variable_comparisons.Rds"))
res_weigh = readRDS(file = file.path("results", "7_weighted_comparisons.Rds"))
res_imput = readRDS(file = file.path("results", "8_rq6_boot_compare_results.Rds"))


```


```{r}
#| include: false

names(res_unadj)
res_unadj$u1cdata1$bootstrap_iter %>% names()

# Data extraction: 
res_unadj$u1cdata1$bootstrap_iter$md #all  mean differences for attritioning on u1cdata1 time point
res_weigh[[1]][[1]]$md # each iteration is stored in different i
res_weigh[[1]][[2]]$md # second iteration
res_imput$u1cdata1$md # first imputed dataset (has 40 bootstrap iterations in each)
res_imput$u1cdata2$md # second imputed dataset

res_unadj$u1cdata1$bootstrap_iter$smd #all  mean differences for attritioning on u1cdata1 time point
res_weigh[[1]][[1]]$smd # each iteration is stored in different i
res_weigh[[1]][[2]]$smd # second iteration
res_imput$u1cdata1$smd # first imputed dataset (has 40 bootstrap iterations in each)
res_imput$u1cdata2$smd # second imputed dataset

res_unadj$u1cdata1$bootstrap_iter$cor_resid #all  mean differences for attritioning on u1cdata1 time point
res_weigh[[1]][[1]]$cor_resid

# LEAVE THE ABOVE FOR NOW - I"LL REMOVE IT LATER.

```

## Create Cleaned Datasets

```{r}
# Correlation pair names (6 unique pairs from 4 variables)
test_correlation_matrix = matrix(
  nrow = length(rq6y),
  ncol = length(rq6y)
)

for (i in seq_along(rq6y)) {
  for (j in seq_along(rq6y)) {
    test_correlation_matrix[i, j] = paste(rq6y[i], rq6y[j], collapse = " ")
  }
}

x = test_correlation_matrix[lower.tri(test_correlation_matrix)]
cor_x_var = str_extract(x, "^\\S+")
cor_y_var = str_extract(x, "\\S+$")
cor_pairs = paste0("cor_", cor_x_var, "_", cor_y_var)

rm(test_correlation_matrix, x)
```

### Unadjusted Dataset

```{r}
# Function to extract and combine unadjusted data for one timepoint
extract_unadj = function(res, timepoint_idx) {
  timepoint = rq1y_twin1[timepoint_idx]
  boot_iter = res[[timepoint]]$bootstrap_iter

  df = tibble(
    timepoint = rq1y_twin_labels_clean_extrashort[timepoint_idx],
    iteration = 1:nrow(boot_iter$md)
  )

  # Add md columns
  md_df = as_tibble(boot_iter$md, .name_repair = "minimal")
  colnames(md_df) = paste0("md_", rq6y)

  # Add smd columns
  smd_df = as_tibble(boot_iter$smd, .name_repair = "minimal")
  colnames(smd_df) = paste0("smd_", rq6y)

  # Add var columns
  var_df = as_tibble(boot_iter$var, .name_repair = "minimal")
  colnames(var_df) = paste0("var_", rq6y)

  # Add cor_resid columns
  cor_df = as_tibble(boot_iter$cor_resid, .name_repair = "minimal")
  colnames(cor_df) = cor_pairs

  bind_cols(df, md_df, smd_df, var_df, cor_df)
}

# Combine all timepoints
df_unadj = map_dfr(seq_along(rq1y_twin1), ~extract_unadj(res_unadj, .x)) %>%
  mutate(condition = "Unadjusted")

df_unadj
```

### Weighted Dataset

```{r}
# Check structure of weighted results
# res_weigh[[timepoint_index]][[iteration]]$parameter

# Function to extract and combine weighted data for one timepoint
extract_weighted = function(res, timepoint_idx) {
  # browser()
  n_iter = length(res[[timepoint_idx]])

  # Pre-allocate lists
  md_list = vector("list", n_iter)
  smd_list = vector("list", n_iter)
  var_list = vector("list", n_iter)
  cor_list = vector("list", n_iter)

  for (i in seq_len(n_iter)) {
    iter_data = res[[timepoint_idx]][[i]]
    md_list[[i]] = iter_data$md
    smd_list[[i]] = iter_data$smd
    var_list[[i]] = iter_data$var
    cor_list[[i]] = iter_data$cor_resid
  }

  # Combine into matrices
  md_mat = do.call(rbind, md_list)
  smd_mat = do.call(rbind, smd_list)
  var_mat = do.call(rbind, var_list)
  cor_mat = do.call(rbind, cor_list)

  # Create tibble
  df = tibble(
    timepoint = rq1y_twin_labels_clean_extrashort[timepoint_idx],
    iteration = 1:n_iter
  )

  # Add md columns
  md_df = as_tibble(md_mat, .name_repair = "minimal")
  colnames(md_df) = paste0("md_", rq6y)

  # Add smd columns
  smd_df = as_tibble(smd_mat, .name_repair = "minimal")
  colnames(smd_df) = paste0("smd_", rq6y)

  # Add var columns
  var_df = as_tibble(var_mat, .name_repair = "minimal")
  colnames(var_df) = paste0("var_", rq6y)

  # Add cor_resid columns
  # BUG FIX: The raw weighted results include 10 correlation values (lower
  # triangle including diagonal). We need only the 6 off-diagonal pairs,
  # at positions 2,3,4,6,7,9 in the 10-element vector.
  # Compute off-diagonal indices dynamically from the number of variables.
  n_vars = length(rq6y)
  dummy_mat = matrix(0, n_vars, n_vars)
  dummy_mat[lower.tri(dummy_mat, diag = TRUE)] = seq_len(n_vars * (n_vars + 1) / 2)
  off_diag_idx = dummy_mat[lower.tri(dummy_mat, diag = FALSE)]
  cor_df = as_tibble(cor_mat, .name_repair = "minimal")[, off_diag_idx]
  colnames(cor_df) = cor_pairs

  bind_cols(df, md_df, smd_df, var_df, cor_df)
}

# Combine all timepoints
df_weigh = map_dfr(seq_along(rq1y_twin1), ~extract_weighted(res_weigh, .x)) %>%
  mutate(condition = "Weighted")

df_weigh
```

### Imputed Dataset

```{r}
# For imputed: 250 imputed datasets × 40 iterations per dataset = 10,000 per timepoint
# res_imput$u1cdata1$md is a list of 40 vectors
# res_imput$u1cdata1, u1cdata2, ..., u1cdata250 for timepoint 1

# Function to extract imputed data for one timepoint
extract_imputed = function(res, timepoint_idx, n_imputed = 250) {
  # Derive prefix from rq1y_twin1 (e.g., "u1cdata1" -> "u1cdata")
  timepoint_prefix = gsub("1$", "", rq1y_twin1[timepoint_idx])

  all_data = vector("list", n_imputed)

  for (m in seq_len(n_imputed)) {
    dataset_name = paste0(timepoint_prefix, m)
    imp_data = res[[dataset_name]]

    n_iter = length(imp_data$md)

    # Extract each parameter
    md_mat = do.call(rbind, imp_data$md)
    smd_mat = do.call(rbind, imp_data$smd)
    var_mat = do.call(rbind, imp_data$var)
    cor_mat = do.call(rbind, imp_data$cor_resid)
    # Create tibble for this imputed dataset
    # Make iteration unique across all imputations: 1-40 for m=1, 41-80 for m=2, etc.
    df = tibble(
      timepoint = rq1y_twin_labels_clean_extrashort[timepoint_idx],
      imputation = m,
      iteration = ((m - 1) * n_iter) + 1:n_iter
    )

    # Add parameter columns
    md_df = as_tibble(md_mat, .name_repair = "minimal")
    colnames(md_df) = paste0("md_", rq6y)

    smd_df = as_tibble(smd_mat, .name_repair = "minimal")
    colnames(smd_df) = paste0("smd_", rq6y)

    var_df = as_tibble(var_mat, .name_repair = "minimal")
    colnames(var_df) = paste0("var_", rq6y)

    cor_df = as_tibble(cor_mat, .name_repair = "minimal")
    colnames(cor_df) = cor_pairs

    all_data[[m]] = bind_cols(df, md_df, smd_df, var_df, cor_df)
  }

  bind_rows(all_data)
}

# Combine all timepoints
df_imput = map_dfr(seq_along(rq1y_twin1), ~extract_imputed(res_imput, .x)) %>%
  mutate(condition = "Imputed")

df_imput 
```

## Verify Datasets

```{r}
#| output: false

head(df_unadj)
head(df_weigh)
head(df_imput)

colnames(df_unadj)
colnames(df_weigh)
colnames(df_imput)

# Check dimensions
cat("Unadjusted dataset:\n")
cat("  Rows:", nrow(df_unadj), "(expected: 30,000 = 10,000 × 3 timepoints)\n")
cat("  Cols:", ncol(df_unadj), "\n\n")

cat("Weighted dataset:\n
")
cat("  Rows:", nrow(df_weigh), "(expected: 30,000 = 10,000 × 3 timepoints)\n")
cat("  Cols:", ncol(df_weigh), "\n\n")

cat("Imputed dataset:\n")
cat("  Rows:", nrow(df_imput), "(expected: 30,000 = 250 × 40 × 3 timepoints)\n")
cat("  Cols:", ncol(df_imput), "\n\n")

# Preview column names
cat("Column names:\n")
names(df_unadj)

```

## checks

```{r}

# gemini, add testthat for all TRUE!
library(testthat) # Ensure testthat is loaded

test_that("Timepoint columns are identical across dataframes", {
  expect_true(identical(df_unadj$timepoint, df_imput$timepoint))
  expect_true(identical(df_unadj$timepoint, df_weigh$timepoint))
  
  df_imput %>%
    filter(timepoint == "Y21") %>% 
    pull(iteration) %>% 
    duplicated() %>% 
    `!` %>%
    all() %>% 
    expect_true()
})

```

# Calculate Bias Reduction

## Data Wrangling

### Main results - bias (not difference)

```{r}
shared_var = Reduce(intersect, list(colnames(df_unadj), colnames(df_weigh), colnames(df_imput)))

combined_results = rbind(df_unadj[shared_var],df_weigh[shared_var],df_imput[shared_var]) %>%  
  select(-starts_with("md_")) %>% 
  pivot_longer(
    cols    = c(-timepoint, -iteration, -condition),
    names_to = "stat"
  )

bias_summary = combined_results %>%
  group_by(condition, timepoint, stat) %>%
  reframe(
    bias = .mean_qi_pd(value)
  ) %>%
  unnest(bias) %>%
  # Split stat into param and variable(s)
  mutate(
    param = str_extract(stat, "^[^_]+"),
    variable_raw = str_remove(stat, "^[^_]+_")
  ) %>%
  # For correlations, extract both variables
  mutate(
    var1 = case_when(
      param == "cor" ~ str_extract(variable_raw, "^[^_]+"),
      TRUE ~ variable_raw
    ),
    var2 = case_when(
      param == "cor" ~ str_extract(variable_raw, "[^_]+$"),
      TRUE ~ NA_character_
    )
  ) %>%
  # Convert to nice labels
  mutate(
    var1_label = rq6y_labels[match(var1, rq6y)],
    var2_label = rq6y_labels[match(var2, rq6y)],
    variable_label = case_when(
      param == "cor" ~ paste0(var1_label, " × ", var2_label),
      TRUE ~ var1_label
    ),
    param = case_when(
      param == "smd" ~ "Standardized Mean Difference",
      param == "var" ~ "Variance % Change",
      param == "cor" ~ "Correlation Change",
      TRUE ~ param
    ),
    param = factor(param, levels = c("Standardized Mean Difference", "Variance % Change", "Correlation Change"))
  ) %>%
  # Order timepoint, condition, and variables
  mutate(
    timepoint = factor(timepoint, levels = rq1y_twin_labels_clean_extrashort),
    condition = factor(condition, levels = c("Unadjusted", "Weighted", "Imputed")),
    var1_order = match(var1, rq6y),
    var2_order = ifelse(is.na(var2), 0, match(var2, rq6y))
  ) %>%
  arrange(param, condition, var1_order, var2_order, timepoint) %>%
  mutate(variable_label = factor(variable_label, levels = unique(variable_label))) %>%
  select(condition, timepoint, param, variable_label, y, ymin, ymax, pval) %>%
  group_by(condition, timepoint, param) %>%
  mutate(
    pval_adj = stats::p.adjust(pval, method = "holm")
  )

bias_summary %>%
  ungroup() %>%
  gt(groupname_col = c("param", "condition"), rowname_col = "variable_label") %>%
  cols_label(
    condition = "Condition",
    timepoint = "Timepoint",
    y = "Est",
    ymin = "LB",
    ymax = "UB",
    pval = md("p"),
    pval_adj = md("p<sub>Holm</sub>")
  ) %>%
  fmt_number(
    columns = c(y, ymin, ymax),
    decimals = 3
  ) %>%
  fmt_number(
    columns = c(pval, pval_adj),
    decimals = 4
  ) %>%
  # Styling - uniform font size
  tab_style(
    style = cell_text(size = px(10)),
    locations = list(
      cells_body(),
      cells_stub(),
      cells_row_groups(),
      cells_column_labels()
    )
  ) %>%
  # Highlight significant positive (green)
  tab_style(
    style = cell_fill(color = "#ffd9d1"),
    locations = cells_body(
      columns = c(y, ymin, ymax, pval, pval_adj),
      rows = pval_adj < 0.05 & y > 0
    )
  ) %>%
  # Highlight significant negative (red)
  tab_style(
    style = cell_fill(color = "#b6c3e0"),
    locations = cells_body(
      columns = c(y, ymin, ymax, pval, pval_adj),
      rows = pval_adj < 0.05 & y < 0
    )
  ) %>%
  # Footnotes
  tab_footnote(
    footnote = md("<em>Note.</em> Est = Estimate, LB = Lower Bound 95% CI, UB = Upper Bound 95% CI. Significant (p<sub>adj</sub> < 0.05) effects highlighted in green (increases) or red (decreases). SMD = (Mean<sub>condition</sub> - Mean<sub>original</sub>) / SD<sub>original</sub>. Variance % change = (Var<sub>condition</sub> - Var<sub>original</sub>) / Var<sub>original</sub>. Correlation change = r<sub>condition</sub> - r<sub>original</sub>."),
    placement = "right"
  ) %>%
  tab_footnote(
    footnote = "P values are Bonferroni-Holm adjusted within each condition × timepoint × parameter group",
    locations = cells_column_labels(columns = pval_adj),
    placement = "right"
  ) %>%
  opt_footnote_marks(marks = c("*", "†"))

```

### Main results - bias reduction relative to unadjusted

```{r}
combined_results_wide = combined_results %>%
  mutate(
    param = str_extract(stat, "^[^_]+"),
    variable_raw = str_remove(stat, "^[^_]+_")
  ) %>%
  select(-stat) %>%
  pivot_wider(
    id_cols = c(timepoint, iteration, param, variable_raw),
    values_from = "value",
    names_from = "condition",
  ) %>%
  ungroup() %>%
  mutate(
    diff1      = Unadjusted - Weighted,
    diffratio1 = (Unadjusted - Weighted)/Unadjusted,
    se1        = (Unadjusted - Weighted)^2,
    diff2      = Unadjusted - Imputed,
    diffratio2 = (Unadjusted - Imputed)/Unadjusted,
    se2        = (Unadjusted - Imputed)^2
  ) %>%
  group_by(param, variable_raw, timepoint) %>%
  reframe(
    pbetter1   = mean(abs(Weighted) < abs(Unadjusted)),
    diff1      = .mean_qi_pd(diff1),
    diffratio1 = .mean_qi_pd(diffratio1),
    mse1       = mean(se1),
    pbetter2   = mean(abs(Imputed) < abs(Unadjusted)),
    diff2      = .mean_qi_pd(diff2),
    diffratio2 = .mean_qi_pd(diffratio2),
    mse2       = mean(se2)
  ) %>%
  unnest(c(diff1, diffratio1, diff2, diffratio2), names_sep = "_")

combined_results_wide


```



#### Bias Reduction: Unadjusted vs Weighted

```{r}
# Prepare diffratio1 table (Unadjusted - Weighted) / Unadjusted
diffratio1_summary = combined_results_wide %>%
  select(param, variable_raw, timepoint, pbetter1,
         y = diffratio1_y, ymin = diffratio1_ymin, ymax = diffratio1_ymax) %>%
  # Add variable labels
  mutate(
    var1 = case_when(
      param == "cor" ~ str_extract(variable_raw, "^[^_]+"),
      TRUE ~ variable_raw
    ),
    var2 = case_when(
      param == "cor" ~ str_extract(variable_raw, "[^_]+$"),
      TRUE ~ NA_character_
    ),
    var1_label = rq6y_labels[match(var1, rq6y)],
    var2_label = rq6y_labels[match(var2, rq6y)],
    variable_label = case_when(
      param == "cor" ~ paste0(var1_label, " × ", var2_label),
      TRUE ~ var1_label
    ),
    param = case_when(
      param == "smd" ~ "Standardized Mean Difference",
      param == "var" ~ "Variance % Change",
      param == "cor" ~ "Correlation Change",
      TRUE ~ param
    ),
    param = factor(param, levels = c("Standardized Mean Difference", "Variance % Change", "Correlation Change")),
    timepoint = factor(timepoint, levels = rq1y_twin_labels_clean_extrashort),
    var1_order = match(var1, rq6y),
    var2_order = ifelse(is.na(var2), 0, match(var2, rq6y))
  ) %>%
  arrange(param, var1_order, var2_order, timepoint) %>%
  mutate(variable_label = factor(variable_label, levels = unique(variable_label))) %>%
  select(param, timepoint, variable_label, y, ymin, ymax, pbetter1) %>%
  mutate(p_val = bayestestR::pd_to_p(pmax(pbetter1, 1 - pbetter1))) %>%
  group_by(param, timepoint) %>%
  mutate(p_adj = p.adjust(p_val, method = "bonferroni")) %>%
  ungroup()

diffratio1_summary %>%
  ungroup() %>%
  gt(groupname_col = "param", rowname_col = "variable_label") %>%
  cols_label(
    timepoint = "Timepoint",
    y = "Est",
    ymin = "LB",
    ymax = "UB",
    pbetter1 = md("P<sub>better</sub>"),
    p_val = md("p"),
    p_adj = md("p<sub>adj</sub>")
  ) %>%
  fmt_number(columns = c(y, ymin, ymax, pbetter1), decimals = 3) %>%
  fmt_number(columns = c(p_val, p_adj), decimals = 3) %>%
  tab_style(
    style = cell_text(size = px(10)),
    locations = list(
      cells_body(), cells_stub(), cells_row_groups(), cells_column_labels()
    )
  ) %>%
  tab_style(
    style = cell_fill(color = "#b6c3e0"),
    locations = cells_body(
      columns = c(y, ymin, ymax, pbetter1, p_val, p_adj),
      rows = p_adj < 0.05 & pbetter1 > 0.5
    )
  ) %>%
  tab_style(
    style = cell_fill(color = "#ffd9d1"),
    locations = cells_body(
      columns = c(y, ymin, ymax, pbetter1, p_val, p_adj),
      rows = p_adj < 0.05 & pbetter1 < 0.5
    )
  ) %>%
  tab_header(
    title = "Bias Reduction: Weighting vs Unadjusted",
    subtitle = md("(Bias<sub>Unadjusted</sub> − Bias<sub>Weighted</sub>) / Bias<sub>Unadjusted</sub>")
  ) %>%
  tab_footnote(
    footnote = md("<em>Note.</em> Est = Estimate, LB = Lower Bound 95% CI, UB = Upper Bound 95% CI. Significant (p<sub>adj</sub> < 0.05) effects highlighted in blue (weighting reduces bias) or red (weighting increases bias)."),
    placement = "right"
  ) %>%
  tab_footnote(
    footnote = md("P<sub>better</sub> = proportion of bootstrap iterations where |Bias<sub>Weighted</sub>| < |Bias<sub>Unadjusted</sub>|. Values > 0.5 indicate weighting reduces bias; values < 0.5 indicate weighting increases bias."),
    locations = cells_column_labels(columns = pbetter1),
    placement = "right"
  ) %>%
  tab_footnote(
    footnote = md("p = two-tailed p-value equivalent derived from P<sub>better</sub> using bayestestR::pd_to_p(). Computed as 2 × min(P<sub>better</sub>, 1 − P<sub>better</sub>)."),
    locations = cells_column_labels(columns = p_val),
    placement = "right"
  ) %>%
  tab_footnote(
    footnote = "p_adj = Bonferroni-corrected p-value within each timepoint.",
    locations = cells_column_labels(columns = p_adj),
    placement = "right"
  ) %>%
  opt_footnote_marks(marks = c("*", "†", "‡", "§"))
```

#### Bias Reduction: Unadjusted vs Imputed

```{r}
# Prepare diffratio2 table (Unadjusted - Imputed) / Unadjusted
diffratio2_summary = combined_results_wide %>%
  select(param, variable_raw, timepoint, pbetter2,
         y = diffratio2_y, ymin = diffratio2_ymin, ymax = diffratio2_ymax) %>%
  # Add variable labels
  mutate(
    var1 = case_when(
      param == "cor" ~ str_extract(variable_raw, "^[^_]+"),
      TRUE ~ variable_raw
    ),
    var2 = case_when(
      param == "cor" ~ str_extract(variable_raw, "[^_]+$"),
      TRUE ~ NA_character_
    ),
    var1_label = rq6y_labels[match(var1, rq6y)],
    var2_label = rq6y_labels[match(var2, rq6y)],
    variable_label = case_when(
      param == "cor" ~ paste0(var1_label, " × ", var2_label),
      TRUE ~ var1_label
    ),
    param = case_when(
      param == "smd" ~ "Standardized Mean Difference",
      param == "var" ~ "Variance % Change",
      param == "cor" ~ "Correlation Change",
      TRUE ~ param
    ),
    param = factor(param, levels = c("Standardized Mean Difference", "Variance % Change", "Correlation Change")),
    timepoint = factor(timepoint, levels = rq1y_twin_labels_clean_extrashort),
    var1_order = match(var1, rq6y),
    var2_order = ifelse(is.na(var2), 0, match(var2, rq6y))
  ) %>%
  arrange(param, var1_order, var2_order, timepoint) %>%
  mutate(variable_label = factor(variable_label, levels = unique(variable_label))) %>%
  select(param, timepoint, variable_label, y, ymin, ymax, pbetter2) %>%
  mutate(p_val = bayestestR::pd_to_p(pmax(pbetter2, 1 - pbetter2))) %>%
  group_by(param, timepoint) %>%
  mutate(p_adj = p.adjust(p_val, method = "bonferroni")) %>%
  ungroup()

diffratio2_summary %>%
  ungroup() %>%
  gt(groupname_col = "param", rowname_col = "variable_label") %>%
  cols_label(
    timepoint = "Timepoint",
    y = "Est",
    ymin = "LB",
    ymax = "UB",
    pbetter2 = md("P<sub>better</sub>"),
    p_val = md("p"),
    p_adj = md("p<sub>adj</sub>")
  ) %>%
  fmt_number(columns = c(y, ymin, ymax, pbetter2), decimals = 3) %>%
  fmt_number(columns = c(p_val, p_adj), decimals = 3) %>%
  tab_style(
    style = cell_text(size = px(10)),
    locations = list(
      cells_body(), cells_stub(), cells_row_groups(), cells_column_labels()
    )
  ) %>%
  tab_style(
    style = cell_fill(color = "#b6c3e0"),
    locations = cells_body(
      columns = c(y, ymin, ymax, pbetter2, p_val, p_adj),
      rows = p_adj < 0.05 & pbetter2 > 0.5
    )
  ) %>%
  tab_style(
    style = cell_fill(color = "#ffd9d1"),
    locations = cells_body(
      columns = c(y, ymin, ymax, pbetter2, p_val, p_adj),
      rows = p_adj < 0.05 & pbetter2 < 0.5
    )
  ) %>%
  tab_header(
    title = "Bias Reduction: Imputed vs Unadjusted",
    subtitle = md("(Bias<sub>Unadjusted</sub> − Bias<sub>Imputed</sub>) / Bias<sub>Unadjusted</sub>")
  ) %>%
  tab_footnote(
    footnote = md("<em>Note.</em> Est = Estimate, LB = Lower Bound 95% CI, UB = Upper Bound 95% CI. Significant (p<sub>adj</sub> < 0.05) effects highlighted in blue (imputation reduces bias) or red (imputation increases bias)."),
    placement = "right"
  ) %>%
  tab_footnote(
    footnote = md("P<sub>better</sub> = proportion of bootstrap iterations where |Bias<sub>Imputed</sub>| < |Bias<sub>Unadjusted</sub>|. Values > 0.5 indicate imputation reduces bias; values < 0.5 indicate imputation increases bias."),
    locations = cells_column_labels(columns = pbetter2),
    placement = "right"
  ) %>%
  tab_footnote(
    footnote = md("p = two-tailed p-value equivalent derived from P<sub>better</sub> using bayestestR::pd_to_p(). Computed as 2 × min(P<sub>better</sub>, 1 − P<sub>better</sub>)."),
    locations = cells_column_labels(columns = p_val),
    placement = "right"
  ) %>%
  tab_footnote(
    footnote = "p_adj = Bonferroni-corrected p-value within each timepoint.",
    locations = cells_column_labels(columns = p_adj),
    placement = "right"
  ) %>%
  opt_footnote_marks(marks = c("*", "†", "‡", "§"))
```

### Mean Squared Error

```{r}

combined_results_mse = combined_results %>%
  mutate(
    condition = factor(condition, levels = c("Unadjusted","Weighted", "Imputed")),
    param = str_extract(stat, "^[^_]+"),
    variable_raw = str_remove(stat, "^[^_]+_"),
    var1 = case_when(
      param == "cor" ~ str_extract(variable_raw, "^[^_]+"),
      TRUE ~ variable_raw
    ),
    var2 = case_when(
      param == "cor" ~ str_extract(variable_raw, "[^_]+$"),
      TRUE ~ NA_character_
    ),
    var1_label = rq6y_labels[match(var1, rq6y)],
    var2_label = rq6y_labels[match(var2, rq6y)],
    variable_pair = paste0(var1_label, " × ", var2_label),
    variable = rq6y_labels[match(variable_raw, rq6y)],
    variable = ifelse(is.na(variable), variable_pair, variable),
    value = value^2
  ) 

combined_results_mse %>% 
  filter(timepoint == "Y26 (web test)") %>% 
  filter(param == "cor") %>% 
  select(-stat) %>% 
  ggplot(aes(y = value, x = condition)) +
  geom_violin(fill = "black") + 
  facet_grid(param ~ variable, scales = "free_y") + 
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90)
  )
  


```


#### Consistently Most Accurate Method (Heatmap)

```{r}
# For each iteration, determine which method has lowest squared error
# Then calculate proportion of times each method wins
mse_consistency = combined_results_mse %>%
  select(variable, param, timepoint, condition, iteration, value) %>%
  pivot_wider(
    names_from = condition,
    values_from = value
  ) %>%
  mutate(
    best = case_when(
      Imputed <= Weighted & Imputed <= Unadjusted ~ "Imputed",
      Weighted <= Imputed & Weighted <= Unadjusted ~ "Weighted",
      TRUE ~ "Unadjusted"
    )
  ) %>%
  group_by(variable, param, timepoint) %>%
  summarise(
    p_unadjusted = mean(best == "Unadjusted"),
    p_weighted = mean(best == "Weighted"),
    p_imputed = mean(best == "Imputed"),
    .groups = "drop"
  ) %>%
  mutate(
    # Only highlight if one method wins >95% of the time
    best_method = case_when(
      p_imputed > 0.95 ~ "Imputed",
      p_weighted > 0.95 ~ "Weighted",
      p_unadjusted > 0.95 ~ "Unadjusted",
      TRUE ~ "None"
    ),
    best_method = factor(best_method, levels = c("Unadjusted", "Weighted", "Imputed", "None"))
  )

# Create heatmap showing consistently winning method
mse_consistency %>%
  mutate(
    param = case_when(
      param == "smd" ~ "SMD",
      param == "var" ~ "Variance",
      param == "cor" ~ "Correlation"
    ),
    param = factor(param, levels = c("SMD", "Variance", "Correlation")),
    timepoint = factor(timepoint, levels = c("Y21", "Y26 (q'aire)", "Y26 (web test)"))
  ) %>%
  ggplot(aes(x = timepoint, y = variable, fill = best_method)) +
  geom_tile(color = "grey50", linewidth = 0.5) +
  facet_wrap(~param, scales = "free_y", ncol = 3) +
  scale_fill_manual(
    values = c("Unadjusted" = "#fee08b", "Weighted" = "#d73027", "Imputed" = "#1a9850", "None" = "white"),
    name = "Consistently Best (>95%)"
  ) +
  labs(
    title = "Most Accurate Method (Lowest MSE)",
    subtitle = "Highlighted only when one method is most accurate in >95% of bootstrap iterations",
    x = NULL,
    y = NULL
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
    axis.text.y = element_text(angle = 45, hjust = 1, size = 7),
    strip.text = element_text(face = "bold", size = 10),
    panel.grid = element_blank(),
    legend.position = "bottom"
  )

save_plot("9_lowest_mse_method", width = 8, height = 8)

```

#### Proportion of Iterations Each Method is Most Accurate (Stacked Bar)

```{r}
# Stacked bar chart showing proportion of times each method wins
mse_consistency_long = mse_consistency %>%
  select(variable, param, timepoint, p_unadjusted, p_weighted, p_imputed) %>%
  pivot_longer(
    cols = starts_with("p_"),
    names_to = "method",
    values_to = "proportion",
    names_prefix = "p_"
  ) %>%
  mutate(
    method = str_to_title(method),
    method = factor(method, levels = c("Unadjusted", "Weighted", "Imputed")),
    param = case_when(
      param == "smd" ~ "SMD",
      param == "var" ~ "Variance",
      param == "cor" ~ "Correlation"
    ),
    param = factor(param, levels = c("SMD", "Variance", "Correlation")),
    timepoint = factor(timepoint, levels = c("Y21", "Y26 (q'aire)", "Y26 (web test)"))
  )

mse_consistency_long %>%
  ggplot(aes(x = variable, y = proportion, fill = method)) +
  geom_col(position = "stack", width = 0.8) +
  facet_grid(timepoint ~ param, scales = "free_x") +
  scale_fill_manual(
    values = c("Unadjusted" = "#fee08b", "Weighted" = "#d73027", "Imputed" = "#1a9850"),
    name = "Method"
  ) +
  scale_y_continuous(labels = scales::percent, breaks = c(0, 0.5, 1)) +
  labs(
    title = "Proportion of Iterations Each Method is Most Accurate (Lowest Mean Squared Error)",
    x = NULL,
    y = NULL
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 20, hjust = 1, size = 8),
    axis.text.y = element_text(size = 7),
    strip.text.x = element_text(face = "bold", size = 9),
    strip.text.y = element_text(angle = 0, size = 8),
    panel.grid = element_blank(),
    legend.position = "right"
  )

save_plot("9_lowest_mse_method_stacked", width = 9, height = 6)

```

# Extra Plots

## Correlation Differences

```{r}

# Shared wrangling used by both the annotations and the main plot
plot_data_cor = combined_results %>%
  mutate(
    condition    = factor(condition, levels = c("Unadjusted", "Weighted", "Imputed")),
    param        = str_extract(stat, "^[^_]+"),
    variable_raw = str_remove(stat, "^[^_]+_")
  ) %>%
  filter(param == "cor") %>%
  mutate(
    var1       = str_extract(variable_raw, "^[^_]+"),
    var2       = str_extract(variable_raw, "[^_]+$"),
    var1_label = rq6y_labels[match(var1, rq6y)],
    var2_label = rq6y_labels[match(var2, rq6y)],
    pair_label = paste0(var1_label, " ×\n", var2_label),
    timepoint  = factor(timepoint, levels = rq1y_twin_labels_clean_extrashort)
  )

# Point estimate + 95% CI per condition × pair × timepoint (for in-panel annotations)
panel_labels = plot_data_cor %>%
  group_by(condition, pair_label, timepoint) %>%
  summarise(
    est = mean(value),
    lo  = quantile(value, 0.025),
    hi  = quantile(value, 0.975),
    .groups = "drop"
  ) %>%
  mutate(label = sprintf("%+.3f [%+.3f, %+.3f]", est, lo, hi))

```

### Descriptive Stats

```{r}

plot_data_cor %>%  
  group_by(condition) %>% 
  summarise(
    .mean_qi_pd(value, .width = .95)
  )

plot_data_cor %>%  
  group_by(condition) %>% 
  summarise(
    .mean_qi_pd(value, .width = .99)
  )

```


### Complex Plot

```{r}

plot_data_cor %>%
  ggplot(aes(x = value, fill = condition, color = condition)) +
  geom_density(alpha = 0.25, linewidth = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey30", linewidth = 0.5) +
  # Annotations: one geom_text per condition so each gets its own colour and vjust row
  geom_text(
    data = filter(panel_labels, condition == "Unadjusted"),
    aes(x = -Inf, y = Inf, label = label),
    color = "#b07800", hjust = -0.05, vjust = 1.3,
    size = 1.8, family = "mono", inherit.aes = FALSE
  ) +
  geom_text(
    data = filter(panel_labels, condition == "Weighted"),
    aes(x = -Inf, y = Inf, label = label),
    color = "#7b241c", hjust = -0.05, vjust = 2.8,
    size = 1.8, family = "mono", inherit.aes = FALSE
  ) +
  geom_text(
    data = filter(panel_labels, condition == "Imputed"),
    aes(x = -Inf, y = Inf, label = label),
    color = "#1a7040", hjust = -0.05, vjust = 4.3,
    size = 1.8big, family = "mono", inherit.aes = FALSE
  ) +
  facet_grid(pair_label ~ timepoint, scales = "free_y") +
  scale_fill_manual(
    values = c("Unadjusted" = "#f0a500", "Weighted" = "#c0392b", "Imputed" = "#27ae60"),
    labels = c("Unadjusted" = "Attritioned (Unadjusted)", "Weighted" = "Attritioned + Weighted", "Imputed" = "Attritioned + Imputed"),
    name = "Method"
  ) +
  scale_color_manual(
    values = c("Unadjusted" = "#b07800", "Weighted" = "#7b241c", "Imputed" = "#1a7040"),
    labels = c("Unadjusted" = "Attritioned (Unadjusted)", "Weighted" = "Attritioned + Weighted", "Imputed" = "Attritioned + Imputed"),
    name = "Method"
  ) +
  labs(
    title    = "Bootstrap Distribution of Correlation Changes by Method and Timepoint",
    subtitle = "Dashed line at zero = no change relative to full-participation sample",
    x        = expression("Correlation change (" * r[attritioned] * " - " * r[original] * ")"),
    y        = "Density"
  ) +
  theme_bw() +
  theme(
    axis.text.x        = element_text(angle = 30, hjust = 1, size = 8),
    axis.text.y        = element_blank(),
    axis.ticks.y       = element_blank(),
    strip.text.x       = element_text(face = "bold", size = 9),
    strip.text.y       = element_text(size = 9, angle = 0, hjust = 0),
    panel.grid.minor   = element_blank(),
    panel.grid.major.y = element_blank(),
    legend.position    = "bottom",
    plot.title         = element_text(size = 11, face = "bold"),
    plot.subtitle      = element_text(size = 9, color = "grey40")
  )

save_plot("9_extra_correlation_densities", width = 8, height = 9)
```

# ACE Estimates

## Extract ACE Differences from All Conditions

### Unadjusted

```{r}
# Unadjusted: use pre-computed bootstrap summaries already stored in res_unadj
ace_diff_unadj = lapply(seq_along(rq1y_twin1), function(i) {
  res_unadj[[rq1y_twin1[i]]]$ace_diff_summary %>%
    mutate(
      timepoint  = rq1y_twin_labels_clean_extrashort[i],
      condition  = "Unadjusted",
      name_clean = rq6y_labels[match(name, rq6y)]
    )
}) %>%
  bind_rows()

ace_diff_unadj
```

### Weighted

```{r}
# Weighted: summarise raw bootstrap iterations
ace_diff_weigh_list = vector("list", length(rq1y_twin1))

for (i in seq_along(rq1y_twin1)) {
  ace_results = lapply(res_weigh[[i]], function(x) x$ace)
  ace_results = lapply(ace_results, function(x) {
    x[[2]]$group = "difference"
    x[[4]]$group = "difference"
    return(x)
  })
  ace_results = lapply(ace_results, function(x) do.call(rbind, x))
  ace_results = do.call(rbind, ace_results) %>%
    filter(par %in% c("a", "c", "e"))

  ace_diff_weigh_list[[i]] = ace_results %>%
    filter(group == "difference") %>%
    group_by(par, name, sex) %>%
    summarise(.mean_qi_pd(value), .groups = "drop") %>%
    mutate(
      timepoint  = rq1y_twin_labels_clean_extrashort[i],
      condition  = "Weighted",
      name_clean = rq6y_labels[match(name, rq6y)]
    )
}

ace_diff_weigh = bind_rows(ace_diff_weigh_list)
ace_diff_weigh
```

### Imputed

```{r}
# Imputed: summarise across bootstrap iterations × imputed datasets
n_imputed_ace = 250

ace_diff_imput_list = vector("list", length(rq1y_twin1))

for (tp in seq_along(rq1y_twin1)) {
  timepoint_prefix = gsub("1$", "", rq1y_twin1[tp])

  all_ace = vector("list", n_imputed_ace)
  for (m in seq_len(n_imputed_ace)) {
    dataset_name = paste0(timepoint_prefix, m)
    imp_data     = res_imput[[dataset_name]]
    # imp_data$ace: list of bootstrap iterations, each a list of dataframes
    ace_iter     = lapply(imp_data$ace, function(x) do.call(rbind, x))
    all_ace[[m]] = do.call(rbind, ace_iter)
  }

  ace_diff_imput_list[[tp]] = bind_rows(all_ace) %>%
    filter(par %in% c("a", "c", "e"), group == "diff") %>%
    group_by(par, name, sex) %>%
    summarise(.mean_qi_pd(value), .groups = "drop") %>%
    mutate(
      timepoint  = rq1y_twin_labels_clean_extrashort[tp],
      condition  = "Imputed",
      name_clean = rq6y_labels[match(name, rq6y_prefix)]
    )
}

ace_diff_imput = bind_rows(ace_diff_imput_list)
ace_diff_imput
```

## Combined ACE Differences Table

```{r}
ace_diff_combined = bind_rows(ace_diff_unadj, ace_diff_weigh, ace_diff_imput) %>%
  mutate(
    condition = factor(condition, levels = c("Unadjusted", "Weighted", "Imputed")),
    timepoint = factor(timepoint, levels = rq1y_twin_labels_clean_extrashort),
    par       = toupper(par),
    sex       = str_to_title(sex)
  ) %>%
  group_by(condition, timepoint, sex) %>%
  mutate(pval_adj = stats::p.adjust(pval, method = "holm")) %>%
  ungroup()

ace_diff_combined %>%
  select(sex, condition, timepoint, par, name_clean, y, ymin, ymax, pval, pval_adj) %>%
  arrange(sex, name_clean, par, condition, timepoint) %>%
  gt(groupname_col = "sex") %>%
  cols_label(
    condition  = "Condition",
    timepoint  = "Timepoint",
    par        = "Par",
    name_clean = "Variable",
    y          = "Est",
    ymin       = "Lower",
    ymax       = "Upper",
    pval       = md("p"),
    pval_adj   = md("p<sub>adj</sub>")
  ) %>%
  fmt_number(columns = c(y, ymin, ymax), decimals = 3) %>%
  fmt_number(columns = c(pval, pval_adj), decimals = 4) %>%
  tab_style(
    style = cell_text(size = px(10)),
    locations = list(
      cells_body(), cells_stub(), cells_row_groups(), cells_column_labels()
    )
  ) %>%
  tab_style(
    style = cell_fill(color = "#ffd9d1"),
    locations = cells_body(
      columns = c(y, ymin, ymax, pval, pval_adj),
      rows = pval_adj < 0.05 & y > 0
    )
  ) %>%
  tab_style(
    style = cell_fill(color = "#b6c3e0"),
    locations = cells_body(
      columns = c(y, ymin, ymax, pval, pval_adj),
      rows = pval_adj < 0.05 & y < 0
    )
  ) %>%
  tab_footnote(
    footnote = md("<em>Note.</em> Difference = Attritioned − Original. Par: A = additive genetic, C = shared environment, E = unique environment. Significant (p<sub>adj</sub> < 0.05) effects highlighted in orange (increases) or blue (decreases). P values Bonferroni-Holm adjusted within each condition × timepoint × sex group."),
    placement = "right"
  )
```

## ACE Bias Reduction Plot

```{r}
ace_diff_combined %>%
  ggplot(aes(
    x     = y,
    xmin  = ymin,
    xmax  = ymax,
    y     = condition,
    color = condition
  )) +
  geom_pointrange(size = 0.3, linewidth = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40", linewidth = 0.4) +
  facet_grid(name_clean + sex ~ par + timepoint, scales = "free_x") +
  scale_color_manual(
    values = c("Unadjusted" = "#f0a500", "Weighted" = "#c0392b", "Imputed" = "#27ae60")
  ) +
  labs(
    title    = "ACE Bias Across Conditions",
    subtitle = "Difference = Attritioned − Original (95% CI). Dashed line = no bias.",
    x        = "ACE Difference",
    y        = NULL,
    color    = "Condition"
  ) +
  theme_bw() +
  theme(
    strip.text.x     = element_text(size = 7, face = "bold"),
    strip.text.y     = element_text(size = 7, angle = 0),
    axis.text.y      = element_text(size = 8),
    axis.text.x      = element_text(size = 7),
    panel.grid.minor = element_blank(),
    legend.position  = "bottom"
  )

save_plot("9_ace_bias_conditions", width = 10, height = 12)
```

